---
title: "Chapter 3: Investigating Speakers and Change Over Time Using Grouped Data"
output: 
  pdf_document:
    number_sections: false
always_allow_html: true
geometry: "left=1in,right=1in,top=2in,bottom=1in"
header-includes:
  - \usepackage{chngcntr}
  - \newcounter{chapter}
  - \setcounter{chapter}{3}
  - \counterwithin{table}{chapter}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

In previous chapters, we investigated how breaking up strings of text
into individual words or multi-word phrases can give us insight into a
corpus. We did things like count the number of words within an entire
data frame. But historians often want to profile not merely the
collective use of language, but also the role of specific speakers and
how their language changed over time.

Understanding change over time begins with the question of what is to be
understood over time. Is it a list of words referring to women? Is it an
individual life? Do we care about the course of change over days, weeks,
years, or decades -- or within a single debate? For many historians, the
answer might well be, "all of the above," which is fine. Counting one
variable (say, references to women) by another variable of time (say,
decades) is how we accomplish an analysis of change over time.

In this chapter, we will learn how to count over time -- and the more
general principle, which is analyzing one variable against another
variable. We may count words by time, or words by speaker to find which
speakers said these words the most.

In the language of the algorithm, understanding the association between
variables is accomplished by "grouping" the data, allowing us to take
other fields, like speaker name, year, and debate title, into account.
This chapter will use these groupings to gain insight into the discourse
of speakers in Parliament.

At this point, critical thinking about what we count is appropriate.The
prospect of analyzing the individual writers in a corpus is an obvious
way to understand the poles of debate. But when the subject of analysis
is the parliament of Great Britain, one understandable reaction may be
skepticism. Why study the language of individual speakers in Parliament
when 19th-century British Parliament was mostly a small group of elite
white men and not people who reflected the British population?

One answer is that parliamentary language is an index of social
experience. We may come to the parliamentary debates interested in the
experience of women or ethnicities; the research program here is neutral
as to what the researcher wants to study.

Issues that rose to parliamentary attention were few. Ordinary people
had to gather petitions to get parliament to entertain the vote for
working people or women. When we look at parliament, we don't see all
the labor that went into assembling the petition -- but we do see a very
important register of when discussions in parliament began to change.
Using parliament as an index of political labor allows us to ask of any
given subject, when did this subject merit institutional attention?
Understanding the flow of power in parliament is a method useful to many
researchers who are otherwise uninterested in identifying the
characteristics of the "great men" of history.

In the exercises that follow, we anticipate that most researchers can
see the value of finding out more about both the men who held powerful
offices and the experience of women. We will therefore touch on both the
mighty -- including prime ministers and chancellors of the exchequer --
and the vulnerable -- showing how we can use data to navigate to the few
women who testified in parliament. Careful navigation of data can teach
us about both kinds of people.

## Counting by Group to Find the Wordiest Speakers

Our first exercise is finding the most "wordy" Parliamentarians from
1830--that is, Parliamentarians who spoke the most words. Often, the
Parliamentarians who spoke the most did so because fellow
Parliamentarians trusted them to express the wants of their own social
group.

To analyze speakers in Hansard, we will first need to import a new
category of data from `hansardr`: `speaker_metadata`.

```{r, message = F}
# load the hansardr library
library("hansardr")

# load the required data 
data("speaker_metadata_1830")

head(speaker_metadata_1830)
```

For now, we will just go over three relevant fields from
`speaker_metadata`.

Each sentence from the Hansard corpus is assigned a unique ID, as
represented by the `sentence_id` field. The speaker who stated a
sentence is assigned the same ID. The shared key across datasets allows
us to join the data from speaker_metadata to other data from `hansardr`,
such as the debate text.

The other two important fields we will address are the `speaker` and the
`suggested_speaker` fields.

The `speaker` field contains the speaker name as it was originally
transcribed within the Hansard corpus. The resulting field contains
inconsistencies and a lack of standardization that can make analysis of
speakers difficult. A single speaker may have been recorded by different
permutations of his first, middle and last name(s) (for example,
"William Gladstone" may be transcribed as "William E. Gladstone", "W.
Gladstone", or "Mr. Gladstone," to name just a few). In other cases, a
speaker may have been called by a rotating office title (like "Prime
Minister").

It is also the case that different speakers are recorded by the same
name. For instance, two people named Sir Robert Peel served in
Parliament during the 1820s and both are evoked by the same spelling of
the name; the older Sir Robert Peel (1750-1830), the industrialist,
briefly overlapped with his son, Sir Robert Peel (1841-46), the future
prime minister. Meanwhile, "Mr. W. Gladstone" could refer to William
Ewart Gladstone or his son William Henry Gladstone.

To make analysis of speaker names more challenging yet, during the
digitization processes, optical character recognition (OCR) errors were
introduced into the speaker names. Common OCR error include interpreting
a lower case "L" as an "i," or the lower case letter "a" as an "o."

Because of these nuances, any naive attempt to count words by speaker
would be guaranteed to produce wrong results. It was paramount,
therefore, that in managing the `hansardr` data, we produce an
authoritative index of actual speakers that corresponded to the facts.
In data science, adding a column with information inking this
authoritative index to the names as presented is called "metadata."

The curators of the dataset -- a team of historians and data scientists
run by the authors -- stored the metadata about the speakers in the
`suggested_speaker` field, building upon the work completed by the
Digging into Linked Parliamentary Data Project and the Egger and
Spirling database (Egger and Spirling, Winters and Steer).

The `suggested_speaker` field contains our suggestion for the true
identity of the speaker. Importantly, the `suggested_speaker` field
represents a marked improvement on existing data sources, tested by
historians, but our data is not guaranteed for universal accuracy; hence
it is only "suggested." In the suggested_speaker field, a unique id is
given that refers to one unique speaker who was in parliament during the
time covered by the database. While multiple speakers during the same
period sometimes share the same name (like the two Gladstones or two
Peels), each individual speaker is assigned a unique number. Thus, when
we see the number 3104 in the suggested_speaker field, the curators of
the database have made the inference that the W. Gladstone in question
refers to William Gladstone, not his son William Henry.

In the following code, we will make our data set smaller and easier to
work with by just selecting the "speaker," "suggested_speaker," and
"text" fields before tokenizing the data into individual words.

```{r, warning=FALSE, message=FALSE}
library("tidytext")
library("tidyverse")
library("lubridate")

data("hansard_1830")
data("speaker_metadata_1830")

words_1830 <- hansard_1830 %>% 
  left_join(speaker_metadata_1830) %>%
  select(speaker, suggested_speaker, text) %>%
  unnest_tokens(word, text) 

head(words_1830)
```

With a data frame containing fields for "speaker", "suggested_speaker",
and "word," we are now in a position to perform a new kind of analysis
to see the number of words each speaker contributed to Parliament using
two new functions: `group_by()` and `summarize()`.

We use `group_by()` to organize the data by group before using summarize
for our count. By using `group_by()` we can count the number of times
one variable occurs, given the presence of another variable. We can
count the number of words spoken by a speaker. We could also use
group_by to count the words spoken in a given year -- or by a given
speaker in a given year, and so on.

The arguments taken by the command `group_by()` tell the computer which
variables to associate with each other. To group by speaker, we make the
argument of `group_by()` speaker, like so:

```{r, message = F}
words_per_speaker_1830 <- words_1830 %>%
  group_by(speaker) 
```

Alternatively, we could group_by suggested_speaker:

```{r, message = F}
words_per_speaker_1830 <- words_1830 %>%
  group_by(suggested_speaker) 
```

A useful function that often accompanies `group_by()` is `summarize()`, 
a command that tells the computer to apply a mathematical transformation
to each of the groups created by group_by. Thus if we group_by(speaker)
and then summarize(), we can count how many words each speaker spoke, or
how many dates a speaker appeared on, or the first date when each
speaker appeared.

The `summarize()` function can be used with any statistical
transformation, for instance `mean()` or `max()`. Here, we will use
`summarize()` with a function to count, which is `n()`. Finally,
`arrange()` allows us to sort the speakers in descending order of total
words spoken.

In the following code the argument `words_spoken = n()` tells
`summarize()` to create a new column with the number of words that
reflect each unique speaker

```{r, message = F}
words_per_speaker_1830 <- words_1830 %>%
  group_by(speaker) %>%
  summarize(words_spoken = n()) %>%
  arrange(desc(words_spoken))

head(words_per_speaker_1830)
```

Most students of British history will be familiar with the names of the
speakers included in the `words_per_speaker_1830` variable, which
include two prime ministers as well as one noted Irish orator and
nationalist, Daniel O'Connell. The count above also shows us that among
the most prolific speakers was the Chancellor of the Exchequer. How
should we interpret this fact?

A naive analyst might assume that the Chancellor of the Exchequer was
one person, the fifth most prolific speaker in the 1830s. Yet this
assumption would be an error of interpretation. The number of words
spoken by the Chancellor counts the speeches of several distinct
individuals.

The careful analyst will be concerned about the speaker classified as
"The Chancellor of the Exchequer," for many individuals held this post
over the course of the years 1830-39, which we are looking at now. The
Chancellor of the Exchequer is the head financial officer of the United
Kingdom.

A more experienced researcher might begin by using the suggested_speaker
field -- not the speaker field -- to count words spoken by each
individual. Notice that the lines of code below differ from the lines of
code above only in that the argument of the function `group_by()` has
been changed from "speaker" to "suggested_speaker:"

```{r, message = F}
words_per_speaker_1830 <- words_1830 %>%
  filter(suggested_speaker != "") %>%
  group_by(suggested_speaker) %>%
  summarize(words_spoken = n()) %>%
  arrange(desc(words_spoken))

head(words_per_speaker_1830)
```

In this view, we see the first and last names of individual speakers as
well as their ID number from parliament (that is, Robert Peel is id
1664; 1664 is not, in this case, a date).

Nevertheless, our first pass revealed something interesting -- as byways
in the archives often do. It is interesting that Hansard represents the
Chancellor of the Exchequer speaking more words than any other office --
including that of Prime Minister -- at least for the 1830s. The
Chancellor of the Exchequer oversees the work of the Treasury. In many
cases, the Chancellor of the Exchequer was a preliminary step to
becoming Prime Minister. Why was the Chancellor of the Exchequer so
important? And what were the various Chancellors talking about that
required so much speechifying?

## A Research Question: What were the Chancellors of the Exchequer talking about?

It may seem arbitrary to investigate the Chancellor of the Exchequer,
but the office provides a perfect opportunity to test how the metadata
indexing individual speakers works. Next, will use the suggested_speaker
field to pull apart the many individuals associated with the office.

Because the dataset includes metadata about each speaker and how they
are described, we can find the many individuals who occupied the office
of the Chancellor of the Exchequer, and we can do research on them
instead. In a sense, it's a happy accident of our research process that
we used `group_by()` applied to the speaker field rather than the
authoritative suggested_speaker field, because the speaker field tells
us about the way that individuals were actually designated in Hansard,
including their office name.

Before we proceed, we should try to refine our research question to
provide maximum historical insight so that we can start to imagine how
our quantitative research process becomes worthwhile. For instance, we
can ask questions about the diversity of views of the individuals who
and we can ask questions about the office.

We would expect most of the Chancellor of the Exchequer's speeches in
parliament to reflect a predominant concern with taxation and spending.
However, it is also possible that different individuals who held this
post prioritized different issues. How different were they, the
Chancellors of the Exchequer for 1830-39?

To analyze the different speakers, the code that follows will use
`group_by()` to group the speeches made by Chancellors of the Exchequer
by suggested_speaker -- the actual individuals behind the office. Then,
we can inspect the number of words used by each individual Chancellor,
comparing their favorite words.

## A Custom Stop Words List

Before we proceed, a first step is to "clean" our data. In previous
chapters we used an existing stop words list from `tidytext`. The
`hansardr` library also provides its own stop words list.

It may be the case, however, that an analyst wants to curate their own
stop words list that caters to their specific data set or research
question. In the following code we create a custom stop words list by
assigning a list of words to a tibble (a `tidyverse`-style data frame).
For our analysis, suppose we want to restrict all words referring to
parliament, ideas, or procedures; we only want to look at words that are
more or less substantive about the subjects being discussed in
parliament.

An appropriate stopwords list might look like this:

```{r}
custom_stop_words = tibble(word = c("hon", "speaker", "house", "question", "lord", "bill", 
"committee", "duty", "country", "time", "amount", "government", "proposed", "law", 
"measure", "learned", "law", "sir", "respect", "public", "gentleman", "gentlemen", 
"friend", "noble", "parliament", "expenditure", "revenue", "tax", "principle", "proposal", 
"consideration", "duties", "stated", "parliament", "act", "opinion", "inquiry", "effect", 
"subject", "object", "motion", "baronet", "crown", "propose", "estimates", "sum", 
"account", "ministers", "majesty", "proposition", "persons", "principles", "service", 
"found", "propositions", "office", "matter", "statement", "paid", "increase", "moved", 
"means", "considerable", "supply", "intention", "debt", "received", "expense", "estimate", 
"charges", "resolution", "notice", "report", "parties", "party", "surplus", "commons", 
"parties", "class", "commission", "form", "answer", "commissioners", "appointment", 
"officers", "saving", "appeared", "granted", "late", "day", "future", "opportunity", 
"saving", "officers", "information", "extent", "authority", "session", "justice", 
"believed", "support", "character", "carried", "plan", "paper", "clause",  "opposite", 
"system", "argument", "rate", "considered", "bills", "increase", "result", "reference", 
"prepared", "laid", "portion", "passed", "intended", "chancellor", "taxation", "classes", 
"appointed", "established", "parish", "confidence", "evidence", "pensions", "bring", 
"income", "connected", "referred", "objection", "hoped", "adopted", "chancellor", 
"account", "ministers", "occasion", "reduction", "reductions", "service", "services", 
"attention", "vote", "brought", "forward", "purpose", "wish", "called", "period", 
"money", "purpose", "power", "charge", "view", "circumstances", "trade", "taxes", "list", 
"civil", "wished", "people", "discussion", "applied", "proper", "taking", "amendment", 
"statements", "repeal", "papers", "fund", "feeling", "measures", "minister", "purposes", 
"payment", "stock", "increased", "bound", "grant", "grounds", "doubt", "applied", "stamp", 
"manner", "conduct", "exchequer", "provide", "reduced", "table", "difficulties", 
"anxious", "compared", "provide", "king", "individual", "expressed", "hear", "hope", 
"establishment", "debate", "contrary", "instance", "introduced", "call", "lords", 
"existing", "half", "head", "offices", "views", "leave", "respecting", "speech", "feel", 
"excise", "effected", "economy", "night", "price", "majority", "reason", "disposed", 
"held", "claims", "leave", "required", "moment", "admitted", "ready", "words", "issue", 
"allowed", "true", "treasury", "administration", "consent", "advantage", "business", 
"calculated", "enter", "fair","nature", "oppose", "reign", "honour", "mode", "reduce", 
"difficulty", "favour", "credit", "adopt", "additional", "individuals", "mind", "sale", 
"times", "amounted", "deficiency", "laws", "til", "agreed", "alluded", "mentioned", 
"meant", "heard", "feelings", "engaged", "consolidated", "conclusion", "circulation", 
"begged", "calculations", "afford", "acts", "acted", "giving", "pledge", "matters", 
"contract", "proceed", "determined", "satisfaction", "declare", "giving", "spirits", 
"resolutions", "pursue", "opposition", "force","pursue", "force", "opinions", "refer", 
"produce", "appeal", "pay", "terms", "private", "reform", "observations", "entitles", 
"laws", "alteration", "due", "gallant", "joint", "undoubtedly", "alluded", "benefit", 
"consequence", "perfectly", "trusted", "loan", "proceed", "cent", "regard", "political", 
"royal", "carry", "importance", "proceedings", "aware", "rates", "similar", "examination", 
"details", "told", "supposed", "sufficient", "person", "necessity", "millions", 
"emoluments", "difference", "department", "till", "submit", "lay", "imposed", "claim", 
"branches", "recommended", "privy", "personal", "select", "salary", "revenues", 
"preceding", "practice", "move", "meet", "hand", "funds", "formed", "expenses", 
"establishments", "entitled", "dimunition", "settlement", "sense", "arrangement", 
"address", "sufficient", "local", "clergy", "relief", "necessity", "application", "claim", 
"responsibility", "situation", "satisfactory", "ground", "revenues", "quarter", "change", 
"founded", "secretary", "provided", "read", "commutation", "body", "provision",  
"difference", "composition", "condition", "hands", "provision", "capital", 
"responsibility", "press", "knowledge", "raised", "friends", "altogether", "meet", "hand", 
"free", "difference", "claim", "necessity", "arguments", "left", "single", "founded", 
"arrangement", "admit", "total", "told", "operation", "apply", "shown", "person", 
"decision", "actual", "returns", "return", "restrictions", "national", "limited", 
"hundred", "existed", "council", "charged", "annual", "allowances", "proceeding", "word", 
"existed", "fairly", "loss", "spirit", "west", "sum"))
```

We are ready to filter and clean our data and explore the langauge of
speakers called Chancellor of the Exchequer.

## Counting the Words of Each Speaker

In the following code we first remove instances of John Spencer and
Rigby Wason, who spoke so few words as Chancellor of the Exchequer that
he contributes little to our project. Then we eliminate numbers, stop
words, and custom stop words before counting each speaker's words with
`group_by()` and `summarize()`.

```{r, message = F}
chancellor_of_the_exchequeur <- words_1830 %>%
  filter(str_detect(speaker, "Exchequer"),
         str_detect(word, "[a-z]"),
         suggested_speaker != "john_spencer_1234",
         suggested_speaker != "rigby_wason_3024",
         suggested_speaker != "",
         suggested_speaker != "chancellor of the exchequerr") %>% 
   anti_join(stop_words) %>%
   anti_join(custom_stop_words) %>%
   group_by(suggested_speaker, word) %>%
   summarize(n = n()) %>%
   top_n(35) %>%
   ungroup()


chancellor_of_the_exchequeur %>% 
  sample_n(5) %>% 
  head()
```

Note that in the table above, the "word" is annotated with information
about the speaker. Annotation is used to prevent confusion about what is
being counted; we're not just counting how many times a word appears,
but how many times each speaker said it.

Next, let's visualize the data so that we can see how many words are
spoken.

```{r, fig.cap="Most frequently used words by each Chancellor of the Exchequer in the 1830s. Bars show the top repeated terms spoken by Henry Goulburn (left panel) and Thomas Spring Rice (right panel). Counts on the x-axis represent the number of times each word appeared in that Chancellor's speeches during the decade.", message = F}

chancellor_of_the_exchequeur <- chancellor_of_the_exchequeur %>%
  mutate(word = reorder_within(word, n, suggested_speaker)) 
  
ggplot(data = chancellor_of_the_exchequeur, 
       aes(x = word, y = n)) +
  geom_col() +
  scale_x_reordered() +
  coord_flip() +
  facet_wrap(~suggested_speaker, scales = "free") +
  labs(x = "word", y = "speaker") +
  ggtitle("Favorite Words of Each Chancellor of the Exchequer in the 1830s")
```

The results show that the individuals who occupied the post of
Chancellor of the Exchequer in fact represented a diversity of
interests, ranging from Goulburn's interest in colonial commodities like
"tobacco" to Rice's interest in the tension between the property and
banking interests. Their territorial interests varied as wel Goulbourn
emphasized England and Ireland, while Rice emphasized Ireland and India

Goulburn is interested in talking about commodities such as beer,
tobacco, and malt; Peel is interested in addressing questions about how
taxes relate to the church and its tithes, as well as political issues
about how dissenters, Protestants and Catholics. Peel also thinks about
commodities, but his commodities are different than Goulburn's; they
include malt, land, and barley. Rice takes the issues in a different
direction, reflecting on Dublin, India, banks, newspapers, and pensions.
He shares with Peel a concern about property or land taxes.

Interpretation always represents an intellectual challenge. A list of
most frequent words is not necessarily an index of importance: the most
frequently-mentioned terms are typically the terms most subject to
debate. Robert Peel, for instance, does not name Ireland's church tithes
because he was interested in protecting the church in Ireland; to the
contrary, he wished to abolish the system of tithes. Nor does the word
list give us material for interpreting the rest of Peel's politics; the
word list does nothing to tell the uninformed analyst about the movement
for the abolition of taxes on grain, which forms a background to
interpreting Peel's mentions of barley. The word list tells us little
about the presence in 1830s Ireland of two churches, Roman Catholic and
the Church of England. Without further reading, we might never known
that Goulburn was famous for arguing for reductions on taxation
(although with this knowledge, we might become interested in the word
"lower"). The analyst who comes to the Hansard corpus without a
background in British history would gain greater insight by reading
secondary sources before interpreting the graph.

After some basic background research, however, the word lists can help
produce insight, cluing the reader into differences such as Goulburn's
relative willingness to talk about "distress," Peel's interest in
"education," or Rice's interest in the post office. With word lists, we
retrieve important clues about the salient differences contributed to
political debates by individual ministers. A researcher interested in
the careers of Chancellors of the Exchequer might start here before
reading more deeply.

There is no rule of thumb for how many words to look at, but more words
is always better. Distant reading, by its nature, is a shortcut to
making sense of long passages of text. We want to take our time with
these words, not rush to conclusions, and not merely read the top five.
If we expanded the list of words, we might get more insight --
especially if we have gone to the trouble of reading more about the work
of Goulborn and Rice during the 1830s and we have specific questions
about their priorities. The researcher is welcome to adjust upwards the
number of words shown in these charts by raising the number in the code
`top_n(35)` -- although at a certain point, legibility will require new
code to expand the bounding box that limits how big a visualization is.

How new is the information we have gleaned to readers of British
history? We have long known that issues of what was taxed and how are a
basic matter for politics of class, empire and nation; everyone wants
someone else to pay for the government. The research modeled in this
textbook has a pedagogical function, and for that reason, we often
present in these chapters examples that could be improved or
investigated further, and the above chart is one of those.

## Critical Thinking About Our Results

In any research process, it is important to reflect on how the choices
we have made impact our analysis, and how making different choices might
have produced different results. For some researchers, the list in
`custom_stop_words` may unnecessarily eliminate words whose usage they
might want to track. Some researchers will want to identify words for
rhetorical statements such as "hoped" or "feeling"; others may be
intrigued about how different speakers refer to "information" or
"knowledge." In general, I have opted for a deep list which includes
every general allusion to governance, institutions, discovery, and
communication. The beauty of custom lists is that each researcher can
easily adjust the words to match their interests Analysts should
understand how carefully they must tailor a custom stop words list to
their project, and how much of a difference additions and subtractions
make. Using a custom stop word list is very much an artisinal skill of
research; it involves much the same skills of judgment, awareness,
curiosity, and sensitivity to background issues from theory and
secondary sources as does careful reading of primary sources.

One complication of stopwords lists is that to produce functional
results, a custom stop words list must often be long; the hundreds of
performative, rhetorical, and governmental words listed below is not a
perfect list, and it would not work for every exercise. In later
chapters, we will investigate approaches such as differential
measurement which allow the reader to skip over custom stop words lists
to obtain information about what distinguishes one speaker from another
or one year from the next. No process is perfect, however, and analysts
must often resort to some kind of custom filter to get closer to
information that is useful for their project. There is no silver bullet
for creating an absolutely accurate tool where text-mining research
always produces useful information; rather, iterative inquiry, paired
with background reading and in-depth reading of primary sources, is the
basis upon which all insight is ultimately made.

## Working With Dates

Knowing who spoke the most in parliament is useful. But very often
insight comes not from examining how counts change over time. What if we
want to know not merely who spoke the most, but who was the top speaker
for each year in parliament? Our `hansard_1830` data frame lists a date
in the "speechdate" column, formatted as a four-digit year, two-digit
month, and two digit day. To easily track speeches by year, we can add a
new "field" -- that is, a column -- listing just the year of each
speech.

Very frequently, when working with information about dates, we need to
extract the month, day, or year from a data set. The `lubridate` package
makes it easy to extract this information from dates in any format.

```{r, message = F}
date1 <- mdy("6/24/1819")
date1

year1 <- year(date1)
year1
```

We will use the `year()` function from `lubridate` with `mutate()` to
create a new column that has just the extracted year from the
`speechdate` field.

```{r, message = F}
library("kableExtra")

### lines 740-780 were not running so I had to add this

data("debate_metadata_1830"); debate_metadata_1830$speechdate <- as.Date(as.integer(unclass(debate_metadata_1830$speechdate)), origin = "1970-01-01")
debate_metadata_1820$speechdate <- as.Date(as.character(debate_metadata_1820$speechdate))
debate_metadata_1830$speechdate <- as.Date(as.character(debate_metadata_1830$speechdate))
debate_metadata_1840$speechdate <- as.Date(as.character(debate_metadata_1840$speechdate))


hansard_1830_w_year <- hansard_1830 %>% 
  left_join(debate_metadata_1830) %>%
  mutate(year = year(speechdate)) 

hansard_1830_w_year %>%
  head() %>% 
  kable()

hansard_1830_w_year %>%
  head() %>%
  mutate(text = str_trunc(text, 120)) %>% 
  kable("html", booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(2, width = "6cm") %>% 
  column_spec(4, width = "4cm")
```

Notice that the resulting dataset has a new field, "year."

If we want to count words by year, first we need to tokenize the
speeches.

```{r, message = F}
words_1830_w_year <- hansard_1830_w_year %>% 
  left_join(speaker_metadata_1830) %>%
  select(speaker, suggested_speaker, text, year) %>%
  unnest_tokens(word, text)

words_1830_w_year %>% 
  head() %>% 
  select(speaker, word, year) %>% 
  kable() #look at the first few columns of the dataset
```

Now that we have a field for the year of each speech, we can return to
"grouping" data to execute a faceted count where we count words by
speaker and year. When we used the command `group_by()` above, we used
it with one argument. But `group_by()` can take multiple arguments,
allowing the analyst to perform grouped counts that involve multiple
dimensions of the dataset. Next, we will use `group_by()` on two data
fields at the same time -- the "speaker" and "year" columns of as the
arguments of `group_by()`, as in `group_by(speaker, year)`.

```{r, message = F}
words_per_speaker_1830 <- words_1830_w_year %>%
  filter(suggested_speaker != "") %>%
  group_by(speaker, year) %>%
  summarize(speaker_words_per_year = n()) %>%
  arrange(desc(speaker_words_per_year))
 
head(words_per_speaker_1830)
```

Notice that the result of grouping and counting is a data set with three
columns: "speaker," "year," and "speaker_words_per_year." The last
column is the count for each speaker per year -- a field that makes it
possible to analyze change over time.

When we use the `group_by()` command, the fields we use for grouping are
preserved in the output. Two columns are the names of the facets we used
for grouping -- "speaker" and "year" -- while the last column,
"words_spoken," is the count of how many words are recorded for each
speaker per year.

### Women in 19th-century Parliament

For the most part, an analysis of speakers in Parliament will take the
researcher down the road of the many elite white men who assumed
powerful positions or passed honorary titles to their sons. Only 12
named women spoke in Parliament throughout the 19th-century. One woman
who spoke the most was Mrs. Walrand from the 1824 debate, "Motion
Respecting the Trial and Condemnation of Missionary Smith at Demerara."
Exploring her role in this debate can give us insight and speculation
into the criteria that had to be set in place for a woman to appear as a
speaker and instrument to the power of Parliament.

"Motion Respecting the Trial and Condemnation of Missionary Smith at
Demerara" reviews the trial, conviction, and death of John Smith, an
Methodist missionary from England assigned to the British slave colony
of Demerara in South America. Mr. Smith's trial accused him of assisting
a slave rebellion that took place in Demerara in August of 1823. MP
Brougham brought the topic of Mr. Smith's trial before Parliament to
argue that the legal proceedings convicting Mr. Smith were unjust and
strayed from the guidance of Britain to the leaders of its Demerara
colony. One could read his testimony as an early foundation for
Britain's movement towards the Slavery Abolition Act of 1833, but his
concerns--as well as the testimonies of other speakers--are nonetheless
intertwined within a biased legal institution.

Mrs. Walrand spoke as a witness to the slave rebellion and served as a
character witness condemning Mr. Smith. According to her testimony, she
saw first hand the brutal violence of the rebels, having been a
"defenceless lady" fired at by these "savages."

```{r, message = F}
data("hansard_1820")
data("speaker_metadata_1820")
data("file_metadata_1820")

mrs_walrand_speaker_metadata <- speaker_metadata_1820 %>%
  filter(str_detect(speaker, regex("Mrs(.*)Walrand", ignore_case = T)))

mrs_walrand_sentences <- left_join(mrs_walrand_speaker_metadata, 
                                   file_metadata_1820, by = "sentence_id") %>%
  left_join(., hansard_1820, by = "sentence_id") %>%
  select(debate_id, text)

mrs_walrand_sentences %>%
  head() %>%
  mutate(text = str_trunc(text, 120)) %>% 
  kable(format = "latex", 
        booktabs = TRUE,
        caption = "Sentences uttered by Mrs. Walrand in the 1820 Hansard debates.") %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(2, width = "10cm") 
```

```{r}
entire_debate <- hansard_1820 %>%
  left_join(file_metadata_1820, by = "sentence_id") %>%
  filter(debate_id == 5142) %>%
  select("text")

debate_table <- tibble(text = head(entire_debate$text, 20))

debate_table %>%
  kable(format = "latex",
        booktabs = TRUE,
        caption = "Excerpt of Debate Text (First 20 Sentences)") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                full_width = FALSE,
                position = "left") %>%
  column_spec(1, width = "16cm")
```

```{r, fig.cap="Most frequently used content words in Mrs. Walrand’s speech. The horizontal bars show the 30 most common words in her speech, ordered by their frequency of use, with counts on the x-axis.", message = F, warning = F}

mrs_walrand_top_words <- left_join(mrs_walrand_speaker_metadata, hansard_1820) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  filter(is.na(as.numeric(word))) %>%
  group_by(speaker, word) %>%
  summarize(n = n()) %>%
  arrange(desc(n)) %>%
  top_n(30)

ggplot(data = mrs_walrand_top_words,
       aes(x = reorder(word, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Mrs. Walrand\'s Top Words",
       subtitle = "From \"Motion Respecting the Trial and Condemnation of 
                   Missionary Smith at Demerara\"",
       x = "Word",
       y = "Count")
```

## Investigating an Event

Next, let's move towards investigating an event.

First, let's organize the data into two groupings -- the words spoken
prereform and those spoken postreform.

```{r, echo=FALSE}
# make two categories, pre- and post-reform
hansard_1820 <- hansard_1820 %>%
  left_join(speaker_metadata_1820) %>%
  left_join(debate_metadata_1820)

hansard_1830 <- hansard_1830 %>%
  left_join(speaker_metadata_1830) %>%
  left_join(debate_metadata_1830)

hansard_1840 <- hansard_1840 %>%
  left_join(speaker_metadata_1840) %>%
  left_join(debate_metadata_1840)

hansard <- bind_rows(hansard_1820, hansard_1830) %>% 
  bind_rows(hansard_1840)

# find all the prereform text and speaker names
prereform <- hansard %>% #hansard_1830 %>%
  select(sentence_id, speechdate, debate, text, speaker, suggested_speaker) %>%
  filter(speechdate <= as.Date("1832-6-4")) %>%
  filter(speechdate >= as.Date("1822-6-4")) %>%
  mutate(year = year(speechdate)) %>%
  mutate(suggested_speaker = coalesce(suggested_speaker, speaker)) %>% # the "coalesce" function replaces a NA in the suggested_speaker column with whatever is in the speaker column
   # filter(!is.na(speaker)) %>% ## a different count will result if we use this line as an alternative to the above
  mutate(period = "prereform")

# find all the postreform text and speaker names
postreform <- hansard %>% #hansard_1830 %>% 
  select(sentence_id, speechdate, debate, text, speaker, suggested_speaker) %>%
  filter(speechdate >= as.Date("1832-6-4")) %>%
  filter(speechdate <= as.Date("1842-6-4")) %>%
  mutate(year = year(speechdate)) %>%
  mutate(suggested_speaker = coalesce(suggested_speaker, speaker)) %>% # the "coalesce" function replaces a NA in the suggested_speaker column with whatever is in the speaker column
  #filter(!is.na(speaker)) %>% ## a different count will result if we use this line as an alternative to the above
  mutate(period = "postreform")
```

Let's break the data up into individual words and count how many words
are in each category.

```{r, echo=FALSE}
# tokenize words (this may be slow -- give it time)
prereform_words <- prereform %>% 
  unnest_tokens(word, text) 
  
postreform_words <- postreform %>% 
  unnest_tokens(word, text) 

# count the words
prereform_word_count <- prereform_words %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  summarise(wordcount = n()) %>%
  ungroup() %>% 
  arrange(desc(wordcount))  %>%
  mutate(period = "prereform")

postreform_word_count <- postreform_words %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  summarise(wordcount = n()) %>%
  ungroup() %>%
  arrange(desc(wordcount)) %>%
  mutate(period = "postreform")

# glue the data together 
both_periods_word_count <- bind_rows(prereform_word_count, postreform_word_count) 

# tell the computer to put the periods in the correct order
both_periods_word_count$period <- factor(both_periods_word_count$period, levels = c("prereform", "postreform"))

# add up the total words for each period
both_periods_word_count_totals <- both_periods_word_count %>%
  group_by(period) %>%
  summarize(total = sum(wordcount))


# create a bar chart
library(scales) # a library useful for formatting numbers

ggplot(both_periods_word_count_totals, aes(x = period, y = total)) +
  geom_col((aes(fill = period))) +
  scale_fill_grey() +
  scale_y_continuous(label=comma) +
  geom_text(aes(label = (paste0(scales::comma(total), " words"))),
            colour = "white",
            vjust = 1.5) +
  labs(title = "How Many Words Were Spoken in Each Period?", x = "", y = "words")

```

Many more words were spoken in the ten years after the reform than the
ten years before. How many more? Text mining means we never have to
guess.

```{r, echo=FALSE}
# subtract the prereform total wordcount from the postreform total word count to find the difference
both_periods_word_count_totals$total[2] - both_periods_word_count_totals$total[1]
```

Next, let's look at the actual words spoken in each period and see how
they compare.

```{r, echo=FALSE}
# top words
top_words <- both_periods_word_count %>% 
  group_by(period) %>%
  slice_max(order_by = wordcount, # put the words in descending order by "wordcount," the column for the count
            n = 10) # the argument "n" for the function "slice_max" takes the number of positions from the top to retain

library(gt)
# make a table
top_words %>%
  gt() %>% # we're using the "great tables" package to make an elegant table
  fmt_number( # format the numbers in the table
    columns = everything(), # apply to all numbers
    decimals = 0, # no decimals
    use_seps = TRUE # use commas
  ) %>%
  tab_header(title = "The Top Ten Words Spoken in Prereform and Postreform Parliament") %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_row_groups()
  ) %>%
  tab_style( # format the column labels
     style = list(
        cell_fill(color = "black"),
        cell_text(weight = "bold", color = "white")
        ),
      locations = cells_column_labels()
     ) %>%
    tab_style( # format the title
      style = list(
        cell_fill(color = "lightgray"),
        cell_text(weight = "bold")
        ),
      locations = cells_title()
      )
```

The analyst will discern that there are only very minor differences
between the top words of both categories. What are we to make of that?

We should begin by admitting that these words tell us little that's new
about parliament or history. The word "house" was frequently used in
parliament to refer to the House of Commons and House of Lords, the two
chambers of parliament. Members of the houses traditionally addressed
each other as "honorable sir" (abbreviated "hon" in Hansard) and (when
addressing peers of the realm) "noble sir," "noble sirs," or "my lord."
The general matters of parliament concerned the state of the "country,"
the passing of "bills" into law, the activities of the "goverment,"
various other "subjects," and the official matter of "questions" -- a
formal category of address about the business of parliament. To find the
top words offers no real insight to any analyst who understands the
basic business of parliament.

The only apparent difference in this table occurs in the words "time" --
a top-ten word prereform but not postreform -- and "Ireland" -- a
top-ten word postreform and prereform. If we looked at more than the
top-ten words, even these differences would likely appear slight; the
two words missing in this table are almost surely in the top twenty list
for both periods -- something we can easily check by adjusting the
command *slice_head(n=10)* from 10 to 20.

In short, this step provides us with what data scientists refer to as a
"validation" -- that is, we have confirmed that the method of counting
words tells us that the speech of parliament concerns the business of
parliament.

This kind of validation step is crucial to knowing that we are on the
right track. If our attempt at counting (or any other statistical
analysis) suggested that parliamentary speech in general concerned
squids, whales, and cephalopods, or even obscure names of towns in
Ireland, we would guess that something had gone wrong. Throughout the
process of data analysis, we should use basic statistical measures as
validation that the general idea is working. If our analysis tells us
anything that doesn't make sense, we should rush to check our method --
not revisit the laws of history.

### Strategies for Reaching Insight

How do we move from validation to insight? There are many strategies.
Let's think through several:

-   

    1)  We can read a longer list of top words for each period, for
        instance the top 500 instead of the top 10. Keeping track of
        minute changes may tell us something. On the other hand, there
        will be a great many words to look at. If the differences are
        slight (like a #9 vs a #12 position for a certain word from one
        period to the next) we risk losing track of which words are
        important for which periods.

-   

    2)  We can look at other aspects of the dataset, for instance the
        speaker column, asking how many words were spoken by each
        speaker in the prereform and postreform periods.

-   

    3)  We can use a variety of statistical methods to aggregate
        information about the speakers that were the most different and
        the words whose expression was most different in the prereform
        and postreform periods.

In the sections that follow, we will undertake each of these steps in
turn. They will give us different answers.

Step #1 is easily accomplished by a quick adjustment in code, tweaking
the command *slice_head(n=10)* from 10 to 500. Please change the command
for youself and inspect the results. Is it easy to interpret them?

Step #2 is easily accomplished by revisiting the code in greater depth.
We won't need any new commands, but we will need to apply the functions
we've already seen -- group_by, summarize, and n() -- to the "speaker"
field rather than the "year" field.

Let's do that now to answer the question, how many words did each
speaker speak in the prereform and postreform period?

## Investigating Speakers' Wordcounts, Pre- and Post-Reform

```{r, echo=FALSE}
# count words per speaker per year 
prereform_speakers_wordcount <- prereform_words %>%
  filter(str_trim(suggested_speaker) != "") %>%
  group_by(suggested_speaker, year) %>%
  summarize(wordcount = n()) %>%
  ungroup() %>%
  mutate(period = "prereform")

postreform_speakers_wordcount <- postreform_words %>%
  filter(str_trim(suggested_speaker) != "") %>%
  group_by(suggested_speaker, year) %>%
  summarize(wordcount = n()) %>%
  ungroup() %>%
  mutate(period = "postreform")

# glue them together
both_periods_speakers <- rbind(prereform_speakers_wordcount, postreform_speakers_wordcount)

# inspect the data
head(both_periods_speakers)
```

Because the team concentrated on speakers who sopke frequently, the data
is much better if we look at the speakers who contributed more than a
few sentences.

```{r, echo=FALSE}
# take the top speakers by wordcount
words_per_speaker <- both_periods_speakers %>%
  group_by(suggested_speaker) %>%
  summarize(wordsperspeaker = sum(wordcount)) 

top_speakers <- words_per_speaker %>%
  slice_max(wordsperspeaker,
            n = 5)

# inspect the data      
top_speakers
```

Here we can see one potential problem -- Lord Henry Brougham appears
both as "Lord Brougham" and as "henry_brougham_1679" when in reality
these are both the same person. The majority of speakers look much
better, though.

## Investigating Speaker Wordcount With Joins

With the data frame both_periods_speakers, we have wordcounts for each
speaker for pre- and postreform. There's a great deal we can do with
this data. First, let's perform some basic counts to help us to
understand, in raw numbers, the basic features of prereform and
postreform parliament.

We'll start by asking how many members of parliament stood for the ten
years pre and post reform. We will label each with a "class" called
"prereform only," "postreform only," or "continuous speakers." Note that
these "classes" are fundamentally different than the "period"
categorization we applied above, where we labeled speeches given before
the 1832 reform law as "prereform" and those after the law as
"postreform." Instead of labeling the years, we want to identify the
speakers that spoke only in one period or the other, and those who spoke
in both.

To identify these speakers, we will begin by returning to an earlier
data frame -- "prereform_speakers_wordcount" -- and using a process
called a "join" to exclude those speakers who also appear in the dataset
"postreform_speakers_wordcount."

Information scientists theorize several ways to combine data through
what is called "set theory." Two of the simplest joins are known as the
"union" of two sets -- what they have in common -- and the "difference"
between two sets -- what is in one but not the other.

To find the prereform-only and postreform-only speakers, we will find
the "difference" between the two sets by using a function named
"anti_join()." The anti_join() function excludes from one dataset all
the data that appear in a second dataset. If we begin with a dataset of
all members of parliament who spoke prereform and perform an "anti-join"
with the set of speakers who spoke postreform, the resulting dataset is
a list of speakers who *only* spoke prereform.

```{r, echo=FALSE}
# Use counting to investigate how many distinct speakers are in the data, pre and postreform

# How many speakers were there prereform?
prereform_only_speakers <- prereform_speakers_wordcount %>%
  select(suggested_speaker) %>%
  anti_join(postreform_speakers_wordcount, by = "suggested_speaker") %>%
  mutate(class = "prereform only") 

# inspect the data
head(prereform_only_speakers)
```

How many speakers are there in our dataset? To answer the question, we
just need to know the number of rows, given that the dataset lists one
distinct speaker for each row. The function *nrow()* will provide an
answer to how many rows there are in any dataset.

```{r, echo=FALSE}
# count the number of speakers in the dataset
nrow(prereform_only_speakers)
```

Performing the same operation again, beginning with postreform speeches
and removing the rows of data that appear prereform, yields a dataset of
speakers who *only* spoke after the reform.

```{r, echo=FALSE}
# How many speakers were there postreform?
postreform_only_speakers <- postreform_speakers_wordcount %>% 
  select(suggested_speaker) %>%
  anti_join(prereform_speakers_wordcount) %>%
  mutate(class = "postreform only") 

# inspect the data
head(postreform_only_speakers)
```

How many speakers were there postreform?

```{r, echo=FALSE}
# count the number of speakers in the dataset
nrow(postreform_only_speakers)
```

We now have two sets of numbers -- the count of unique speakers who
spoke 1822-32 and 1832-42. These numbers are new to history; there has
never before been a count of these speakers. We can learn something from
them. Before text mining, it was impossible to get a real impression of
how many speakers were active in parliament before and after 1832. But
the simple ability to count allows us to add detail to our accounts of
the reform and its mechanisms.

The basic finding is that there were more speakers after the reform
period. This makes sense, because the reforms of 1832 introduced more
popular consent into the democratic process. No longer were members
appointed in "pocket buroughs" by aristocrats who held the sole
authority for electing the representative. Now, all seats were open to a
popular vote (albeit of those rich enough to pay at least £10 per year
in taxes -- no mean sum). If we know that the number of seats in the
House of Commons were set at 658 before and after the reform, the
numbers tell us that this reform introduced more variation and turnover
in who occupied the seats of parliament.

Substantively more different members of parliament were elected in the
ten years following 1832 than in the decade before. *How many more
members were there among the postreform only members than the prereform
only members?*

```{r, echo=FALSE}
nrow(postreform_only_speakers) - nrow(prereform_only_speakers)
```

We can also investigate speakers from a third category -- those who were
continuously in parliament, who held office before the reform and who
were able to secure election after the reform as well. How many such
speakers were there?

Please note that we do not have the data to ask which members of
parliament who held office before 1832 were *elected* to office after
1832. We do not have the data to support that question. What we have in
Hansard is only a record of who was recorded as having *spoken.* This is
a rather more discrete query, because not every elected member actually
speaks when they are in office, and some of those who spoke might have
spoken so little that the clerks recording speeches took no notice of
what they said.

Because recorded parliamentary speech is an index of importance, the
members who were recorded as having spoken before and after 1832 offers
a list of those members who were important enough to have taken notice
of the public in both periods. This is a good proxy for which members
mattered in the public imagination *and* met the criterion for having
been elected both before and after the reform. Do we imagine this number
to have been large or small?

We might guess that it was vanishingly small, because a reform of
historical significance could have swept away all incumbents. On the
other hand, many historians of the 1832 Reform Bill have pointed out
that it didn't shift the election process so drastically as had once
been assumed.

In any case, to identify the speakers who appear in both the prereform
and postreform datasets, we will use another function -- inner_join() --
to identify what is known as the "intersection" of the two datasets.
That is, we are looking for speakers who occupy a row in both the
*prereform_speakers_wordcount* data frame and the
*postreform_speakers_wordcount* data frame. The syntax for inner_join()
is just like that of anti_join(), but the result is to *keep* rows that
appear in both datasets, not to exclude those rows.

```{r, echo=FALSE}
# How Many Members of Parliament Spoke Both Prereform and Postreform?
continuous_speakers <- prereform_speakers_wordcount %>%
  distinct(suggested_speaker) %>%
  inner_join(postreform_speakers_wordcount) %>%
  ungroup() %>%
  mutate(class = "both periods") %>%
  distinct(suggested_speaker, class)

nrow(continuous_speakers)
```

We now have actual numbers for the members of parliament recorded as
only having spoken before the reform, those who were recorded as only
having spoken after the reform, and those who were recorded as having
spoken both before and after the reform. Let's represent our results
visually.

```{r, echo=FALSE}
# Count the number of speakers in each category

# glue together the data from each smaller dataset
all_classes <- rbind(prereform_only_speakers, postreform_only_speakers, continuous_speakers)

# group by period and count the number of speakers
all_classes_count <- all_classes %>%
  group_by(class) %>%
  summarize(speakercount = n()) %>%
  arrange(desc(class))

# tell the computer to put the periods in the correct order
all_classes_count$class <- factor(all_classes_count$class, levels = c("prereform only", "postreform only", "both periods"))

# create a bar chart
ggplot(all_classes_count, aes(x = class, y = speakercount)) +
  geom_col((aes(fill = class))) +
  scale_fill_grey() +
  geom_text(aes(label = paste0(speakercount, " speakers")),
            colour = "white",
            vjust = 1.5) +
  labs(title = "How Many Speakers are in Each Category?", x = "", y = "speakers")
```

Certain caveats are important for interpreting this number. While the
speaker data gives us reliable insights into relative trends in
membership and speech, the numbers are only so good as our dataset.

One of the great advantages of text mining is that we can analyze not
just the speakers in parliament, but the consequences of the 1832 Reform
Act for what was spoken about in parliament. Let's begin by asking what
parliament spoke about before and after the reform.

## Working With Word Count

```{r, echo=FALSE}
## Using Word Count: 
## How much did speakers speak before and after the reform act?

# glue the data together for processing
all_per_speaker_wordcounts <- bind_rows(prereform_speakers_wordcount, postreform_speakers_wordcount) %>%
  ungroup() 

# create a database of speaker, year, and wordcount
all_speakers_annotated <- all_per_speaker_wordcounts %>%
  left_join(all_classes) %>%
  filter(!is.na(suggested_speaker)) %>%
  ungroup() 

# inspect the data
all_speakers_annotated
```

Note that we have created two columns that annotate the speeches and
speakers in different ways. The "period" column refers to the year in
question -- were the n words said in 1831 prereform or postreform? The
"class" column sorts the speakers themselves -- did this speaker only
speak in the prereform years, the postreform years, or in both periods?

We also have a *wordcount* column that reflects the words each speaker
spoke per year.

Next, let's sort the data to find the speakers who spoke the most per
year within each category -- prereform, postreform, and continuously
through the periods.

To do this, we'll want to first create a new column,
*wordcount_per_speaker*, which adds together each speaker's total
wordcount. If we have multiple "wordcount" columns, we should add
information about what each refers to. To make it less confusing, we'll
rename our previous *wordcount* column *wordcount_per_year*. It's good
practice to keep renaming fields in our datasets for clarity so that we
don't get confused about what they mean. To calculate
*wordcount_per_speaker*, we'll use the *group_by()* and *mutate()*
functions that we've seen before. Note that we won't use *summarize()*
in this case because we still want to keep the *wordcount_per_year()*
field upon which the new count is based.

we'll use *group_by()* to group the data by "class," then use the
*arrange()* and *slice_head()* functions to take the top wordcounts for
each category.

```{r, echo=FALSE}

# find the total wordcount_per_speaker
wordcount_per_speaker <-  all_speakers_annotated %>%
  rename(wordcount_per_speaker_per_year = wordcount) %>% 
  group_by(class, suggested_speaker) %>%
  summarize(wordcount_per_speaker = sum(wordcount_per_speaker_per_year)) 

# find the most prolific n speakers of each period
top_speakers <- wordcount_per_speaker %>%
  filter(!is.na(suggested_speaker)) %>%
  group_by(class) %>%
  arrange(desc(wordcount_per_speaker)) %>%
  slice_head(n=5) %>%
  mutate(top = TRUE)

#inspect the data
top_speakers
```

## Annotating Our Data

Before we go any further, let's do some work to make the speaker names
more elegant. Because of the structure of our data, names that were
algorithmically cleaned are labeled with a standardized name and number,
e.g. "robert_peel_1665." When we make a graph to share, we'd like to
substitute one of Robert Peel's names as given in the text for this
awkward string. The names as given in the text are in the *hansard*
dataset under the *speaker* column. We can grab any one of them -- minus
the titles like "Lord Chancellor" with which some speakers are
referenced -- and it will read like a normal name. The *str_detect()*
function is useful for searching for a string. In this case,
*!str_detect()* will allow us to throw out names that include titles we
don't want to use. The *sample_n()* function is useful here as it tells
the computer to select a random pick from a set; in other words, if
we've grouped Robert Peel's many possible names, and we have the set
"Robert Peel," "Sir Robert Peel," and "Mr. Robert Peel," the computer is
instructed to pull out one of these at random.

After we choose one name to use from the full Hansard list of literal
speaker references, we can use these names to annotate the previous
dataset, *all_speakers_annotated*, which has the *wordcount* (per
speaker per year) and speaker class.

Let's try it.

```{r, echo=FALSE}
# clean up speaker names
speaker_key <- rbind(prereform, postreform) %>%
  select(speaker, suggested_speaker) %>%
  filter(!str_detect(speaker, "Secretary")) %>%
  filter(!str_detect(speaker, "Chancellor")) %>%
  filter(!str_detect(speaker, "Minister")) %>%
  group_by(suggested_speaker) %>%
  sample_n(1)

# inspect the data
speaker_key %>% filter(str_detect(suggested_speaker, "peel_")) 

```

The results may not be impressive for the vast majority of speakers who
spoke very little. But for major speakers like Robert Peel, this
*speaker_key* dataset will help us reconcile the many names with which
they appeared in the text with one numeric id number, then reconciling
that numeric id with one recognizable name from the Hansard
transcription.

# Using Left Join to Annotate the Data

We've previously used set theory and "joins" to find the intersection
and difference between two datasets. Here, we'll use another kind of
join -- the "union" -- to annotate one dataset with information from
another dataset.

That is, in *all_speakers_annotated*, we have a list of speakers and how
many words they spoke every year. If we use *left_join()* to merge that
dataset with the dataset *top_speakers()*, we'll have a record of
whether each speaker was one of the most prolific speakers in their
category. If we use *left_join()* to merge the dataset again with the
dataset *speakers_key()*, we'll be able to use the literal
transcriptions of speaker names to annotate the visualization later.

The function *left_join()* takes as its first argument the name of the
dataset with which we're merging our data. A second optional argument is
the "by" argument, which allows us to specify a column or set of columns
to use for matching. In the case of the first join, we'll specify two
fields, concatenated into a list, using the syntax *by =
c("suggested_speaker", "class").*

```{r, echo=FALSE}
speakers_w_top_and_speaker_key <- all_speakers_annotated %>%
  rename(wordcount_per_speaker_per_year = wordcount) %>%
  left_join(top_speakers, by = c("suggested_speaker", "class")) %>%
  left_join(speaker_key) %>%
  distinct()

speakers_w_top_and_speaker_key %>% 
  filter(str_detect(suggested_speaker, "peel_")) %>% 
  select(speaker, top, year, wordcount_per_speaker_per_year, class) %>%
  kable()
```

# Visualize the Results

Sometimes the most effective way of analyzing data is to "inspect" the
results of counting, as we have done henceforth. For simple results,
like a list of names or words, we prefer a table.

But for questions about information when several dimensions of data are
related to each other -- for instance speaker words over time when
speakers have been categorized in three "classes" -- visualization may
be useful for conveying how many factors are related to each other.
Next, we'll use *ggplot()* to create a visualization of the results as a
timeline.

In the case of the data about speakers before and after 1832, we have
rich information that deserves to be represented in an interesting way.
We have three categories of speakers -- prereform, postreform, and
continuous speakers -- and information about how many speakers are in
each category. We also have information about how much each speaker
spoke each year. Can we model this information together and learn
something from the result?

in the code block below, we'll create a "dot plot," mapping one point
for each of the 10,810 speakers in the data frame. We will use the color
and shape of the dots to map the three classes of speakers -- prereform,
postreform, and continuous.

### Contingent labeling

Because labeling all the speakers would result in an unreadable graphic,
we will tell the computer to only label those speakers who are
classified as "TRUE" in the column *top*, where we annotated if a
speaker was among the top n speakers of each class by wordcount. The
syntax for this operation is given inside the parentheses of
*geom_label_repel()*, where the coder has an opportunity to point to a
new dataset as the source of the text in the labels. Here, we tell the
computer which dots to label and which to leave blank with the use of
*subset(graph_ready, top == TRUE)*. Operations of this type are
relatively advanced for visualization, but they are extremely useful for
the kind of dense, text-forward information that implementors of text
mining often want from their analysis.

If you are new to visualization with *ggplot()*, we encourage you to
play around with the code below to understand how the components work.
Try changing the size of the dots by adjusting the "size" parameter
inside *geom_jitter()*. Or try changing from a *geom_jitter()* plot to a
*geom_point()* plot. For an in-depth tutorial about the powerful
*ggplot()* library and the philosophy of the "grammar of graphics"
behind it, we encourage you to consult Hadley Wickham, *A Layered
Grammar of Graphics,"* Journal of Computational and Graphical
Statistics,\* vol. 19, no. 1, pp. 3–28, 2010.

```{r, echo=FALSE}
# graph it 
library(ggrepel) # load a useful package for labeling a dense plot

ggplot(speakers_w_top_and_speaker_key, aes(x = year, y = wordcount_per_speaker_per_year, color = class, shape = class)) +
  geom_jitter( # it's worth replacing the geom_jitter with geom_point here and in the next line to see the stakes of diff representations
    size = 3
    ) +
  scale_color_grey(guide = guide_legend()) +
  geom_label_repel(data=subset(speakers_w_top_and_speaker_key, top == TRUE),
                      aes(label=speaker, color = class)) +
  guides(fill = "none", label = "none", size = "none") + 
  theme(legend.position="bottom") +
  labs(title = "How Much Did Speakers Speak", 
  subtitle = "Before and After the Second Reform Act of 1832?") 
```

What do we learn from this plot? A great deal. One is about the
relatively higher counts of speeches in the 1830s than the 1820s,
irrespective of the reform Act. Speeches by lord Althorp, Robert Peel,
and Henry Brougham regularly reached into range of 150,000 to 200,000
words -- which if delivered at breakneck speed would have taken a
stunning *fifteen hours* to deliver.

Even the speeches of lesser speakers were longer in the new decade,
stretching towards 50,000 words instead of 20,000 words as the upper
limit for most rank-and-file speakers -- meaning that after the reform,
multiple speakers each year addressed the house for *five hours* or more
at a time.

We also learn that by and large it was the continuous speakers -- not
the prereform-only aristocratic appointees or the postreform-only
members elected by the middle class -- who spoke at the greatest length.
Represented with black dots, these men included prime ministers like
Robert Peel.

We can also use this visualization to note the few prereform-only
speakers who delivered relatively long speeches, for instance Mr.
Canning and Mr. Huskisson, the latter important to the Board of Trade,
represented by light gray squares.

Equally exceptional were the few members of parliament elected only
after 1832 who spoke with prominence, among them John Roebuck and
Frederick Shaw. These men whose careers were endorsed by reform swam in
a sea of seasoned representatives who had first joined parliament in an
earlier era.

Note that our visualization is only as good as the quality of the data.
Where speakers have not been reconciled correctly -- as is the case with
"Lord Brougham" (classified as postreform only), who in reality is the
same as "Henry Brougham" (classified as a continuous speaker) -- they
may appear as discrete speakers when indeed they are the same.

# Exercises

## 1) As we discussed in the beginning of this chapter, analyzing the individual or collective speeches of men in Parliament can give insight into the workings of the Parliament as an institution of power. We can explore the roles of elite white men as they exerted their will--and the will of their peers--through the Parliamentary institution. We can also explore the roles of women in Parliament and examine how women differ in their contributions, as they were prohibited from becoming members of Parliament. Further, the woman's presence had to be mediated by one of the MPs, as they could not be there without being selected or approved. In result, women were treated as instruments of power in Parliament.

We can further explore how the role of women manifested in Parliament.
First, count the total number of words women spoke in Parliament. Refer
to the table below, which visualizes the women in Parliament and the
year in which they spoke. Can you find a pattern across the kinds of
topics women are brought to Parliament to speak on?

```{r, echo = FALSE, message = FALSE}
name <- c("Mrs. (Mary Ann) Clark", "Mrs. Bridgeman", "Miss Mary Ann Taylor", "Louisa Demont",
          "Franchette Martigner", "Mrs. Walrand")

name_col_2 <- c("Ms. Cunninghame Graham", "Mrs. Dillon", "Mrs. M'Govern", 
                "Mrs. Mitchel-Thomson", "Ms. Cave", "")

year <- c("1809", "1809", "1809", "1820", "1820", "1824")

year_col_2 <- c("1887", "1902", "1902", "1907", "1907", "")

df <- data.frame(name, year, name_col_2, year_col_2)

df %>%
  kable(col.names = c("Name", "Year", "Name", "Year")) %>%
  kable_styling()
```

Note: The way in which their names are recorded in this table might not
reflect the way in which their names appear in the debates because the
names have been consolidated into one representation

```{r womenwords, echo = FALSE}

data("hansard_1800"); data("speaker_metadata_1800")
data("hansard_1820"); data("speaker_metadata_1820")
data("hansard_1880"); data("speaker_metadata_1880")
data("hansard_1900"); data("speaker_metadata_1900")
# Bind text data
all_text <- bind_rows(
  "1800" = hansard_1800 %>% select(sentence_id, text),
  "1820" = hansard_1820 %>% select(sentence_id, text),
  "1880" = hansard_1880 %>% select(sentence_id, text),
  "1900" = hansard_1900 %>% select(sentence_id, text),
  .id = "dataset"
)

# Bind metadata
all_metadata <- bind_rows(
  "1800" = speaker_metadata_1800, 
  "1820" = speaker_metadata_1820, 
  "1880" = speaker_metadata_1880, 
  "1900" = speaker_metadata_1900, 
  .id = "dataset"
)

# Define the list of women's names 
women_patterns <- c(
  "Mrs\\.?\\s*(\\(Mary\\s*Ann\\)\\s*)?Mary\\s*Ann\\s*Clark(e)?", 
  "Mrs\\.?\\s*Bridgeman",
  "Miss\\s*Mary\\s*Ann\\s*Taylor",
  "Louisa\\s*Demont",
  "Franchette\\s*.*Martign\\w+",          
  "Mrs\\.?\\s*Walrand",
  "Ms\\.?\\s*Cunninghame\\s*Graham",
  "Mrs\\.?\\s*Dillon",
  "Mrs\\.?\\s*M['’]?Govern",
  "Mrs\\.?\\s*Mitchell?\\s*[- ]?Thomson",  
  "Ms\\.?\\s*Cave"
)

women_regex <- str_c(women_patterns, collapse = "|")
women_speech <- all_metadata %>%
  filter(str_detect(speaker, regex(women_regex, ignore_case = TRUE))) %>%
  mutate(speaker = str_remove(speaker, regex("^Then\\s+", ignore_case = TRUE))) %>%
  left_join(all_text, by = c("sentence_id", "dataset")) %>%
  select(speaker, text, dataset) %>%
  filter(!is.na(text)) %>%
  unnest_tokens(word, text)

# Count Total Words per women
total_words_by_woman <- women_speech %>%
  count(speaker, sort = TRUE)

print("Total Words Spoken by Each Woman:")
print(total_words_by_woman)

# total of ALL women
total_words_women <- women_speech %>% summarize(total_words = n())
total_words_women

```

```{r wometopic, echo = FALSE}
# Identify Topics 
data("stop_words")

women_topics <- women_speech %>%
  anti_join(stop_words, by = "word") %>%
 filter(!str_detect(word, "^\\d+$")) %>%
  count(dataset, speaker, word, sort = TRUE) %>%
  group_by(dataset, speaker) %>%
  slice_max(n, n = 10, with_ties = FALSE) %>%
  ungroup()

print("Top Topics (Words) for Each Woman:")
print(women_topics)
```

The data shows us a clear pattern, in that women in the 19th century
were almost exclusively brought to Parliament as witnesses or subjects
of inquiry rather than as independent political actors like men. Mrs.
Bridgeman was the clear top woman speaker with over 8000 words spoken,
using terms like "evidence," "question," "examined". Following as
second, Mrs. Walrand's use of words like "insurrection," "slaves,"
"martial" show us that these women were most likely providing testimony
on specific events, crimes, or colonial crises rather than debating
policy as MPS. As we saw earlier in the chapter, these women functioned
as "instruments of power." Their voices were mediated by male MPs, who
selected them primarily to substantiate or refute specific legal claims.

## 2) Use the techniques shown above to move between a distant reading of the top words spoken by Mrs. Walrand or Louisa Demont in all of her speeches to the full debate. (As a reminder, you can view all of a speaker's speeches by filtering for just that speaker. You will also need to load the data for the appropriate years.) How does transitioning between the different modes of reading text provide insight into this specific debate? How does the data frame illuminate the role of women in 19th-century Parliament?

```{r walrand, echo = FALSE}
data("hansard_1820")
data("speaker_metadata_1820")
data("file_metadata_1820")

# from chapter 3 code
mrs_walrand_speaker_metadata <- speaker_metadata_1820 %>%
  filter(str_detect(speaker, regex("Mrs(.*)Walrand", ignore_case = TRUE)))

# Distant reading
mrs_walrand_top_words <- mrs_walrand_speaker_metadata %>%
  left_join(hansard_1820, by = "sentence_id") %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words, by = "word") %>%
  filter(!str_detect(word, "^\\d+$")) %>%
  count(word, sort = TRUE)

mrs_walrand_top_words %>% 
  slice_head(n = 30)
```

```{r walrandsent, echo = FALSE}

mrs_walrand_sentences <- mrs_walrand_speaker_metadata %>%
  left_join(file_metadata_1820, by = "sentence_id") %>%
  left_join(hansard_1820, by = "sentence_id") %>%
  select(debate_id, sentence_id, speaker, text)

mrs_walrand_sentences %>% 
  head(20)

walrand_debate_id <- mrs_walrand_sentences %>%
  distinct(debate_id) %>%
  slice(1) %>%
  pull(debate_id)

entire_debate <- hansard_1820 %>%
  left_join(file_metadata_1820, by = "sentence_id") %>%
  filter(debate_id == walrand_debate_id) %>%
  arrange(sentence_id) %>%
  select(sentence_id, text)

entire_debate %>% 
  slice_head(n = 20)

```

Moving from distant to close reading shifts the question from what Mrs.
Walrand talks about to how her words function in the debate. When
looking just at the keywords and their counts in distant reading, we see
top terms like "smith" "house" "circumstances" which implies discussion
of a trial. But seeing the actual debate text reveals she was painting a
"horror narrative" rather than debating policy or a trial. In our close
reading of the full sentences, we can see that the word "Smith" is in
reference to a Parson Smith, who is described in the debate as a ‘spiritual master and director’ of enslaved people,and the testimony uses racialized language that frames him as culpable for unrest among the enslaved population. Walrand uses emotional imagery of a "dying, bleeding man" in
"excruciating agony" which functions to build sympathy for white settlers. Most
importantly, we see that this testimony is actually part of a male MP's
speech - "Whatever may be the credit attached to this latter testimony,
I quote it to shew the crimes" - confirming that women were not present
as independent legislators, but used as character witnesses to lend
emotional credibility to the establishment’s defense of the white
population.

## 3) Adjust the above code to search for a different office title. What can you infer about the different MPs who held this title? Can we

support our inferences through close readings of their speeches? To get
you started, here are a few office positions: - Chancellor of the
Exchequer - Prime Minister - Attorney General - Lord Chancellor

```{r office, echo = FALSE}

data("debate_metadata_1830")

office_title <- "Lord Chancellor"

office_tokens <- words_1830 %>%
  filter(
    str_detect(speaker, regex(paste0("^\\s*(the\\s+)?", office_title, "\\b"), ignore_case = TRUE)),
    !is.na(suggested_speaker),
    suggested_speaker != "",
    str_detect(word, "^[a-z]+$")
  ) %>%
  anti_join(stop_words, by = "word")

office_words <- office_tokens %>%
  count(suggested_speaker, word, sort = TRUE) %>%
  group_by(suggested_speaker) %>%
  mutate(prop = n / sum(n)) %>%
  slice_max(prop, n = 15, with_ties = FALSE) %>%
  ungroup() %>%
  left_join(speaker_key, by = "suggested_speaker") %>%
  mutate(word = reorder_within(word, prop, speaker))

```

```{r officeplot, fig.width=12, fig.height=7, message=FALSE, echo = FALSE}
office_words_plot <- office_words %>%
  mutate(speaker_id = str_remove(suggested_speaker, "_\\d+$")) %>%
  mutate(chancellor = case_when(
    speaker_id == "henry_brougham" ~ "Henry Brougham (Lord Chancellor)",
    speaker_id == "john_copley"    ~ "John Copley - Lord Lyndhurst (Lord Chancellor)",
    speaker_id == "charles_pepys"  ~ "Charles Pepys - Lord Cottenham (Lord Chancellor)",
    TRUE ~ speaker
  )) %>%
  mutate(word = reorder_within(word, prop, chancellor))

ggplot(office_words_plot, aes(x = word, y = prop)) +
  geom_col() +
  scale_x_reordered() +
  coord_flip() +
  facet_wrap(~ chancellor, scales = "free_y", ncol = 2) +
  labs(
    title = str_c("Favorite Words of Each ", office_title, " in the 1830s"),
    x = "word",
    y = "share of tokens"
  )
```

Analyzing the office of the Lord Chancellor reveals a distinct
linguistic profile compared to that of the Chancellor of the Exchequer.
The top words for Lord Chancellors like Henry Brougham and John Copley
include terms like "noble," "bill," "law," and "court," which confirms
that the office's main function was as the highest legal authority in
the land, mediating procedural and judicial matters in the House of
Lords. Unlike the Exchequer's focus on financial terms like "malt,"
"tax," and "revenue," the Lord Chancellor's vocabulary is much more
institutional and procedural, reflecting their dual role as a judge and
a presiding officer.

## 4) At the beginning of this chapter, we compared counts are based on the "speaker" column with counts based on the "suggested_speaker"

column. Expand each table to show 30 rows. How are the two tables
different? What are some pros and cons about visualizing the original
speaker names and the disambiguated speaker names? Identify a future
research question based on what you learn.

```{r, echo = FALSE}
# speaker
words_1830 %>%
  group_by(speaker) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  slice_head(n = 30)

# suggested_speaker 
words_1830 %>%
  filter(suggested_speaker != "") %>%
  group_by(suggested_speaker) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  slice_head(n = 30)
```

The speaker table shows us individuals by their specific titles - such
as separating "Mr. Stanley" (272,226 words) from "Lord Stanley" (235,627
words) - while the suggested_speaker table correctly consolidates them
with ID, under edward_stanley_2326 (603,787 words). We also see that the
Speaker list is cluttered with office titles like The Chancellor of the
Exchequer whereas the suggested list correctly attributes those speeches
to the individuals holding the office. Methodologically, this is the
distant and close-reading lesson applied to metadata - speaker reflects
the archive’s naming practices and social hierarchy at the moment of
utterance, while suggested_speaker is the better unit for longitudinal
claims about individual participation - with the caveat that
consolidation depends on the quality of the disambiguation algorithm.

## 5) Find the speakers who were referred to by the greatest variety of names in the 1830s. Share the results. Investigate; what characteristics do you think they had in common? AI’s are great for

tasks involving linked data — that is , background about individuals —
especially for the individuals who are well known and who had published
obituaries and biographies, like many members of parliament. Feed the
resulting list to an AI and ask it for input about what they had in
common. What did you learn?

```{r variety, echo = FALSE}
name_variety <- speaker_metadata_1830 %>%
  filter(!is.na(suggested_speaker), suggested_speaker != "") %>%
  group_by(suggested_speaker) %>%
  summarise(
    unique_names = n_distinct(speaker),
    aliases = paste(sort(unique(speaker)), collapse = "; "),
    .groups = "drop"
  ) %>%
  arrange(desc(unique_names))

name_variety %>% slice_head(n = 30)

```

The AI investigation revealed that the speakers with the greatest name
variety share a common characteristic: they are high-status, front-bench
politicians. Figures like Robert Peel and Henry Brougham appear under
multiple aliases because Hansard labels them both by personal name and
by office (for example, “Mr./Sir/Lord X” versus “The Lord Chancellor”),
and those labels can shift with promotions, knighthoods, peerages, or
changes in post during the 1830s. Their high volume of participation
also increases the chance of inconsistent transcription, punctuation,
and spelling variants in the record, which further multiplies
“different” speaker strings for the same person. Taken together, this
shows why raw speaker names can fragment and undercount the influence of
the most powerful individuals in Parliament, making the disambiguated
suggested_speaker column essential for accurate person-level analysis -
while still leaving open a separate research question about how changes
in titles themselves signaled authority over time.

## 6) Find the speakers who held the highest number of offices in the 1830s. Share the results. Then, investigate with AI: what do these

individuals have in common? Tell us what you learned.

```{r holders, echo = FALSE}
office_holders <- words_1830 %>%
  filter(!is.na(suggested_speaker), suggested_speaker != "") %>%
  filter(str_detect(suggested_speaker, "_\\d+$")) %>%
  filter(!str_detect(suggested_speaker, "\\|")) %>%  
  filter(str_detect(speaker, "^The\\s")) %>%
  filter(!str_detect(speaker, regex("^The\\s+(Duke|Earl|Marquess|Viscount|Baron|Bishop)\\b", ignore_case = TRUE))) %>%
  group_by(suggested_speaker) %>%
  summarise(
    office_count = n_distinct(speaker),
    offices = paste(sort(unique(speaker)), collapse = "; "),
    .groups = "drop"
  ) %>%
  arrange(desc(office_count))

cat("Speakers who held the highest number of distinct offices in the 1830s:\n")
office_holders %>%
  slice_head(n = 30) %>%
  print(n = 30)

```

In my results, the top “multi-office” speakers are overwhelmingly senior
front-bench figures - especially major legal officers and cabinet-level
administrators - whose roles were salient enough that Hansard frequently
indexed them by office title as well as by personal name. The
investigation suggests a clear shared profile among the leaders in
office count, such as John Campbell (11 titles), Charles Pepys, and
Thomas Spring Rice (7 titles), as senior legal and administrative
officials experiencing rapid career mobility. In particular, legal
figures like Campbell and Pepys moved through a recognizable
nineteenth-century *cursus honorum*, cycling through posts such as
Solicitor General to Attorney General to Lord Chancellor, with each
transition generating new labels that Hansard records as distinct
speaker strings.

Methodologically, this shows us that the “office count” metric is a
composite of real political advancement and metadata fragmentation.
These individuals appear as “multi-office” speakers partly because they
accumulated offices, but also because they were prominent enough to be
addressed interchangeably by name and by their changing titles. The
practical takeaway is that `suggested_speaker` is essential for stable
person-level analysis, while office-based `speaker` labels are useful
for studying how institutional authority is signaled in the record. A
close-reading follow-up would be to compare a debate where one of these
figures appears under an office title versus one where they appear under
a personal name, since we would expect different rhetorical styles when
speaking in an official capacity (procedural, authoritative,
justificatory) vs. as a private member.

## 7) We’ve used a list of speakers’ words over time to investigate many aspects of the data, but let’s go further. For the continuous speakers, almost everyone talks more after the reform. But for whom is the difference how much they spoke pre reform and post reform the greatest? Share your results. Again, use AI to investigate what those speakers have in common. Tell us what you learned.

```{r growth, echo = FALSE}

pre_totals <- prereform_speakers_wordcount %>%
  filter(!is.na(suggested_speaker), str_trim(suggested_speaker) != "") %>%
  group_by(suggested_speaker) %>%
  summarise(total_pre = sum(wordcount), .groups = "drop")

post_totals <- postreform_speakers_wordcount %>%
  filter(!is.na(suggested_speaker), str_trim(suggested_speaker) != "") %>%
  group_by(suggested_speaker) %>%
  summarise(total_post = sum(wordcount), .groups = "drop")

speech_growth <- pre_totals %>%
  inner_join(post_totals, by = "suggested_speaker") %>%
  mutate(word_diff = total_post - total_pre) %>%
  arrange(desc(word_diff))

print("Continuous speakers with the greatest PRE vs POST difference:")
print(head(speech_growth, 10))
```
The speakers with the largest PRE to POST increase in speech volume - John Russell, Robert Peel, Thomas
Spring Rice and Daniel O’Connell share the role of parliamentary managers in the 1830s. Their explosion in word count reflects a shift in how Parliament functioned after 1832. Compared to the pre-Reform era, the post-1832 House operated more like a “working legislature.” It confronted massive, complex bills, such as the Slavery Abolition, the New Poor Law, and Irish Coercion, which required sustained, procedural, line-by-line management rather than occasional set-piece oratory. Russell and Lord Althorp were leaders responsible for steering government business, and they would have needed to constantly intervene in order to introduce clauses, address objections, and manage amendments. Peel’s growing prominence as a party leader similarly demanded frequent responses and strategic positioning, while O’Connell’s increased volume aligns with Ireland’s significance as a crisis of the decade. This stemmed from his role as a mass-movement parliamentary leader, which sparked repeated debates, petitions, and confrontations that expanded his speaking footprint. AI’s analysis of the growth pattern suggests that, aside from adding new members, the reform intensified the workload and conflicts within leadership, shifting Parliament from a forum for speeches to a machine for legislation.
Methodologically, the key caution is that pre/post word totals can mix real behavioral change with archival
effects (reporting density, office-label conventions, or exits from the Commons), so a close-reading follow-up would compare one pre- and post-1832 debate for a top-growth speaker to confirm whether the increase
reflects more procedural interventions, a longer, more justified framing, or a change in rhetorical strategy.

## 8) Use your code to identify the continuous speakers who have the greatest difference between the year during which they spoke the most and the year during which they spoke the least. Share your results. Again, use AI to investigate what those speakers have in common. Tell us what you learned.

```{r contspeakers, echo = FALSE}
continuous_gap <- all_speakers_annotated %>%
  filter(class == "both periods") %>%
  group_by(suggested_speaker) %>%
  summarise(
    max_words = max(wordcount, na.rm = TRUE),
    min_words = min(wordcount, na.rm = TRUE),
    gap = max_words - min_words,
    .groups = "drop"
  ) %>%
  arrange(desc(gap))

continuous_gap %>% slice_head(n = 20)
```



The speakers with the greatest difference between their maximum and minimum yearly word counts - such as Henry Brougham, John Spencer (Lord
Althorp), and Robert Peel - are not randomly inconsistent. Their
volatility is structural. The massive gaps usually correspond to years when they held major government offices (for example, Chancellor of the Exchequer, Leader of the House, or other senior ministerial posts), which required constant interventions to introduce clauses, answer objections, manage amendments, and keep legislation moving, versus years when they were out of office, in opposition, ill, traveling, or otherwise less central to the agenda. For example, John Spencer’s extreme gap is consistent with his role as Leader of the House during the Reform crisis: in peak sessions he would have had to speak repeatedly simply to steer business, while in quieter years his participation drops sharply. More broadly, the 1830s include intense,
time-bounded conflicts (reform, Ireland, administrative and fiscal
restructuring), so a figure can show bursts concentrated in narrow
windows rather than steady participation across the decade.

What I learned is that “most talkative” is not a stable personality
trait - it often reflects political responsibility and crisis
conditions. Methodologically, this warns against using a single year of word counts as a proxy for long-term influence, because speech volume is
frequently dictated by the office held at that moment and by
event-driven peaks in parliamentary workload. It also cautions against
interpreting one-year snapshots in downstream text analysis (keywords, topic models) without context, since a “trend” may simply be the
minister’s busy year rather than a durable ideological shift. A useful
follow-up would be to check whether each spike-year coincides with holding a specific office, and then rerun the comparison within a single debate topic or smooth across multiple years to separate event-driven surges from longer-term patterns.

## 9) Rework the final visualization so that what is labeled are not the top speakers overall, but the top speakers in the prereform and

postreform categories. Show your results.

```{r visual, echo = FALSE, fig.width = 10, fig.height = 5, message = FALSE, warning = FALSE}
top_speakers_pr_post <- wordcount_per_speaker %>%
  filter(class %in% c("prereform only", "postreform only")) %>%
  group_by(class) %>%
  slice_max(wordcount_per_speaker, n = 5, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(top = TRUE)

graph_ready_pr_post <- all_speakers_annotated %>%
  filter(class %in% c("prereform only", "postreform only")) %>%
  rename(wordcount_per_speaker_per_year = wordcount) %>%
  left_join(
    top_speakers_pr_post %>% select(suggested_speaker, class, top),
    by = c("suggested_speaker", "class")
  ) %>%
  mutate(top = dplyr::coalesce(top, FALSE)) %>%
  left_join(speaker_key %>% distinct(suggested_speaker, .keep_all = TRUE),
          by = "suggested_speaker") %>%
  mutate(
    clean_name = speaker %>%
  str_remove_all("\\b(The|Mr|Sir|Lord|Viscount|Earl|Duke|Marquess)\\.?\\b") %>%
  str_remove("^\\.+\\s*") %>%
  str_remove(regex("^of\\s+", ignore_case = TRUE)) %>%
  str_squish()

  ) %>%
  distinct()

label_df <- graph_ready_pr_post %>%
  filter(top) %>%
  group_by(class, suggested_speaker) %>%
  slice_max(wordcount_per_speaker_per_year, n = 1, with_ties = FALSE) %>%
  ungroup()

ggplot(graph_ready_pr_post, 
       aes(x = year, y = wordcount_per_speaker_per_year, group = suggested_speaker)) +
  
  geom_line(
    data = graph_ready_pr_post %>% filter(!top),
    color = "grey80", linewidth = 0.3, alpha = 0.5
  ) +
  
  geom_line(
    data = graph_ready_pr_post %>% filter(top),
    aes(color = clean_name),
    linewidth = 1.2, alpha = 0.9
  ) +
  
  geom_point(
    data = graph_ready_pr_post %>% filter(top),
    aes(color = clean_name),
    size = 2
  ) +

geom_text_repel(
  data = label_df,
  aes(label = clean_name, color = clean_name),
  size = 3,
  fontface = "bold",
  box.padding = 0.35,
  point.padding = 0.2,
  min.segment.length = 0,
  show.legend = FALSE,
  seed = 340
) +
  
  facet_wrap(~ class, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 12, face = "bold", color = "black"),
    plot.title = element_text(face = "bold")
  ) +
  labs(
    title = "Who Dominated the Debate?",
    subtitle = "Top speakers labeled within prereform-only\nvs postreform-only cohorts",
    x = "Year",
    y = "Words spoken (per speaker-year)"
  )
```

This visualization compares which “prereform only” vs “postreform only” speakers dominated speech volume over time, highlighting different elite cohorts before and after reform.

## AI Use Statement

I used AI tools in the following ways:

**Tool:** Chatgpt

**Purpose:** I used
ChatGPT to identify historical background on the MPs (e.g., cabinet
roles, legal offices, party leadership) after generating the tables for Questions 5–8 in R to support interpretations.

I also used Chatgpt to troubleshoot. I encountered memory/vector allocation issues ("vec_size") when
running the official chapter code for joining the 1820, 1830, and 1840
datasets. I used ChatGPT to help troubleshoot this error, which resulted
in breaking the loading process into separate chunks to get it to run.

**Example Prompts:** 
“I’m working with Historic Hansard 1830s data. Here is a
list of suggested_speaker IDs with the most distinct raw speaker
aliases. What do these MPs have in common historically?”

“Error in left_join: Can’t join on sentence_id because of incompatible
types - what does this mean and how can I check types?”

**Independent Work:** I wrote the R code for cleaning, regex filtering,
counting, and visualization myself, and I wrote all historical claims in
my own words based on the outputs and close-reading.
