---
title: "Chapter 4: Text Mining British Concepts in the 1833 Debates About the Abolition of Slavery"
output: 
  pdf_document:
    number_sections: false
geometry: "left=1in,right=1in,top=2in,bottom=1in"
header-includes:
  - \usepackage{chngcntr}
  - \newcounter{chapter}
  - \setcounter{chapter}{4}
  - \counterwithin{table}{chapter}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The British Abolition of Slavery in 1833, alongside the abolition of the
slave trade by British naval vessels in 1806, were once celebrated as
the sign that Britain was a nation committed to moral progress around
the world. That claim has been significantly modified by historical
research in recent years, as researchers have underscored that the
abolition of slavery failed to end many forms of slavery; it wooed slave
owners through instituting a prolonged form of slavery called
"apprenticeships." In India and the rest of Asia, slavery meanwhile
persisted unreformed.

Studying the interplay between white male politicians as they debated
the fate of millions of enslaved persons from the West Indies to South
Africa is in many ways an out-of-date exercise. Today, we recognize the
degree to which enslaved people across British empire struggled to
liberate themselves. A persistent tradition of uprisings and riots
marked their ongoing resistance. The life of ideas and leaders — even
women leaders like Queen Nannie — are important to understanding the
British Empire.

We will find only an indirect trace of how the freedom struggles of the
Caribbean changed laws and ideas in Britain. The personages and
uprisings that rocked the West Indies were rarely named in Westminster.
The direct records of those events are held in pamphlets and documents
in the Caribbean itself, slowly being digitalized through other means.

And yet, those freedom struggles half a world a way did matter in many
respects to British life. They circulated among evangelical readers;
they forced philosophers of capitalism to reckon with the way that
capitalism had been instantiated as misery in many places where it was
enacted. British debates about slavery came to a head in the halls of
parliament after 1832, when an evangelical middle class got the vote
after decades of political struggles of their own. In 1833, the elected
representatives of the evangelical middle class committed themselves to
talking about slavery as a moral ill and bringing it to an end across
empire.

One question that might allow us to unpack the impact of abolition is
how British understandings of empire and capitalism were changed in the
course of those debates — in a year when both parliament and the
newspapers were filled with a detailed discussion of the suffering
caused by the slave system. One way of approaching the question of what
happened in 1833 is to look for the ideas that were up for debate. Were
any of the fundamental concepts of the enlightenment challenged or
redefined in the course of these arguments?

# The Importance of Background Knowledge

Before we begin to think about the contribution that text mining can
make to the problem, we first need to be able to summarize the
understanding and debates of the historians who have studied these
questions. The paragraphs above summarize some of the different
perspectives on the slavery question taken by historians from the recent
and more distant past. In short, we need some background knowledge. But
how does an analyst get there, if they don't first have a series of
prepared paragraphs to introduce them to a historical problem?

"Background knowledge" in a subject field refers to the kind of
knowledge that one acquires by reading the major textbooks and secondary
monographs from the field of History that cover the historical time
period in question. Textbooks and secondary monographs provide extensive
knowledge about which sources have been consulted, what scholars have
debated, whether the actions of kings or parliaments or working-class
people mattered at this period of time and why.

Remarks about the importance of reading background information may be
experienced as off-putting by some undergraduate students of data, who
want to be assured that in a course on text mining that they will
acquire the tools to make wonderful discoveries. But just as students of
statistics must study more than an introductory course in statistics to
make a novel contribution, so too students of text mining must enroll in
more than one course to do really important work. Reading a wikipedia
entry cannot substitute for the kind of background knowledge that a
student would acquire in the course of taking an introductory course on
the History of Victorian England. Nor can Text Mining for Historical
Analysis make up for courses in the undergraduate study of History,
although this book’s companion volume, The Dangerous Art of Text Mining,
includes many arguments about the specific kind of knowledge of sources,
obfuscations, agency, and experience that historical study makes
available. Producing insight worthy of the attention of multiple fields
is no simple game. It cannot be produced at the touch of a button or the
application of a new algorithm to new data; it requires adjusting the
algorithm, rethinking the questions, examining the data, and iterating
through the work until something truly surprising has come to light.

A shortcut which Guldi recommends in The Dangerous Art of Text Mining
for those who want to achieve results before enrolling in a specialist
course on the historical background material is to engage in an ongoing
collaboration with a professor, librarian, or graduate student already
familiar with the background material, allowing that scholar’s sense of
validation and surprise to guide the work with data. In ‘hybrid' work,
where computer scientists and humanists work on the same team to produce
new knowledge, collaborating over multiple iterations and
investigations, there can be a division of labor around the research. In
general, a high caliber of work can only be accomplished by teams where
at least one valued member is as serious about history as they are about
algorithmic study.

Not every historian needs access to every kind of background knowledge.
If a course on Victorian Britain offers the ideal background for text
mining the parliamentary debates, a course on American history offers a
much better background to working with the Congressional Record.
Scholars who have studied Latin American, African, or Asian history will
necessarily have the background to work with other datasets. The point
here is that understanding a time and a place represents a commitment of
reading dozens of books. No single Wikipedia article or textbook will
substitute for this work, and only this background work really equips an
analyst to ask the questions about consensus and surprise that outfit an
analyst for understanding when their data work is validating what
scholars already know and when their data work has advanced to the
discovery of something truly surprising.

For the purpose of our question about the abolition debatees in
parliament, we recommend extra background reading in the major sources
we consulted to summarize the debates above:

-   David Ryden, West Indian Slavery and British Abolition, 1783-1807
    (Cambridge: Cambridge University Press, 2005)
-   Adam Hochschild, Bury the Chains: Prophets and Rebels in the Fight
    to Free an Empire (Boston: Houghton Mifflen, 2006).
-   Christopher Leslie Brown, Moral Capital: Foundations of British
    Abolitionism (Chapel Hill: University of North Carolina Press, 2006)
-   Richard Huzzey, Freedom Burning: Salvery and Empire in Victorian
    Britain (Ithaca: Cornell University Press, 2012).
-   Claudius K. Fergus, Revolutionary Emancipation, Slavery and
    Abolitionism in the British West Indies (Baton Rouge, Louisiana
    State University Press, 2013).
-   Robert Burroughs and Richard Huzzey, eds., The Suppression of the
    Atlantic Slave Trade (Manchester: Manchester University Press, 2015)
-   Paula Dumas, Proslavery Britain: Fighting for Slavery in an Era of
    Abolition (New York: Palgrave Macmillan, 2016).
-   Padraic X. Scanlan, Slave Empire: How Slavery Built Modern Britain
    (Boston: Little, Brown, 2020).
-   Tom Zoellner, Island on Fire: The Revolt that Ended Slavery in the
    British Empire (Cambridge: Harvard University Press, 2022).

Note that we have attempted to gather titles from the past two decades,
with perspectives ranging from an overview of the abolition movement to
the biography of the abolitionists, to the pro-slavery lobby that argued
against abolition and for apprenticeships, to the larger question of how
imbricated slavery was within the British economy, to the Jamaican
revolt of 1831 that forced Britain to end slavery.

Engagement with historical secondary sources is one of the markers of
serious Digital History work which is frequently absent in publications
from Computer Science or Data Science, with the result that it is
sometimes unclear what the original contribution of a given article is
to the historical literature. We believe that setting a higher standard
of innovation in scholarship –- by demanding findings that are literally
surprising to the field -- is a healthy practice, raising the bar for
computational investigations of the human past.

The analyst beginning research on a problem of this kind does not
necessarily have to read all of these books with equal attentiveness.
But it is useful to spend a few hours reading at least the introductions
to these books (or more recent contributions to the debate) to see what
the authors agree and disagree about, which themes or ideas they find
most interesting.

In the sections of text that follow, as we pursue the question more
deeply, we will return to secondary sources, seeking out more
specialized texts, for instance biographies of the individuals whose
words we count. Pay attention to the footnotes and the texts that we
cite to understand how the need for background knowledge evolves with
research.

# Finding the Research Question that the Tools of Text Mining Can Address

First, we need to think about the fit between the tools of text mining
and the issue of the abolition of slavery. As explained above, we don't
have the source base in Hansard to contribute to burning questions about
the role of enslaved persons in pursuing their own freedom. If all we
have is Hansard, we need to retailor our question to the dataset we
have. With Hansard, we can count mentions of slavery and the way that
slavery is addressed by the spakers who were in power at the time when
slavery was nominally abolished throughout British empire. Can we write
a research question that might light to insights on the basis of word
count?

One place to start is with the speakers in parliament who contributed
the most to the 1833 debates about slavery. As we saw in a previous
chapter, one place where text mining excels is simply counting the
number of words contributed to a given issue by individual speakers.
Based on word counts, we can expect to accurately identify the speakers
in parliament who contributed the most to the slavery question.

We can then move to asking about the language they shared to the
language that makes each of their contributions unique. Another place
where text mining excels is counting the words from each individual
speaker, finding the words they shared and the words that were most
distinctive of each speaker.

Why would we want to understand the words that made the principal
speakers different? Because it may tell us something important about
what they disagreed about. An understanding of the poles of debate can
lead us to a detailed case about the tensions that emerged in parliament
around abolition.

From this question we may be able to navigate to an original
contribution about the role of arguments in parliament, perhaps
contributing information that other historians have not been able to
pull from their research.

## Top Speakers in the Debates About Slavery

In a previous chapter, we looked into the top speakers in parliament in
a given year. In this exercise, we want to begin by finding just those
speakers who spoke the most in debates about slavery. We can use
`str_detect()` to look for the word "slavery" in the titles of debate.
Then we could the top speakers.

If you are new to code, take a moment to notice how changing only one
line in the code below produces a different angle of analysis. The code
below is otherwise very similar to the code in the previous chapters.
One of the great pleasures of coding is the facility with which a few
skills can quickly become powerful in the hands of an analyst who has
many questions.

```{r, warning=FALSE, message=FALSE, cache=FALSE}
# load packages
library("tidyverse")
library("tidytext")
library("lubridate")
library("hansardr")
library("ggrepel")

# load data
# load data
data("hansard_1830")
data("debate_metadata_1830")
data("speaker_metadata_1830")

# merge the Hansard speech text with metadata (e.g., debate title and speech date)
# create a new column 'year' by extracting the year from the full speech date
# find just the debates about slavery and who spoke them 
# FIX: repair the corrupt IDate *before* any left_join()
debate_metadata_1830$speechdate <- as.Date(
  as.integer(round(as.numeric(debate_metadata_1830$speechdate))),
  origin = "1970-01-01"
)

slavery_1830 <- hansard_1830 %>%
  left_join(debate_metadata_1830, by = "sentence_id") %>%
  filter(str_detect(str_to_lower(debate), "slavery")) %>%
  mutate(year = lubridate::year(speechdate))
# This creates a stable year variable so we can study how attention to slavery shifts across the 1830s.Extract and add the year from the 'speechdate' column

slavery_debates_1830 <- slavery_1830 %>%
  left_join(speaker_metadata_1830, by = "sentence_id") %>% # join datasets 
  select(sentence_id, speaker, suggested_speaker, text, year) %>% # keep relevant columns 
  filter(suggested_speaker != "") # remove any lines with missing speaker information

# break up the text into words
slavery_words_1830 <- slavery_debates_1830 %>% # create a new dataset
    unnest_tokens(word, text) # break up the text into words

# count the words per speaker
words_per_speaker_slavery_1830 <- slavery_words_1830 %>% # create a new dataset
  group_by(speaker) %>% # group the data by speaker and year 
  summarize(words_per_speaker = n()) %>% # create a column with total words per speaker
  arrange(desc(words_per_speaker)) # arrange the words in descending order 

# find the top speakers 
top_slavery_speakers <- words_per_speaker_slavery_1830 %>% # create a new dataset
  ungroup() %>% # remove the previous data grouping 
  arrange(desc(words_per_speaker)) %>% # arrange in descending order 
  slice(1:20) # take only the top n speakers

# calculates the number of words each speaker said per year in the 1830 slavery dataset
# group the text data by speaker and year, count the words, 
# and arrange the results in descending order
words_per_speaker_per_year_slavery_1830 <- slavery_words_1830 %>% # create a new dataset
  group_by(speaker, year) %>% # group by speaker and year
  summarize(words_pspy = n()) %>% # count words and store in new column
  arrange(desc(words_pspy)) # sort in descending order of word count

# create a summary of the number of words each speaker contributed per year 
# in the slavery debates of 1830
# the resulting dataset helps identify the most vocal speakers 
# and their yearly distribution of speech.
words_per_speaker_per_year_slavery_1830 <- slavery_words_1830 %>% # create a new dataset
  group_by(speaker, year) %>% # group by speaker and year
  summarize(words_pspy = n()) %>% # count words per speaker per year
  arrange(desc(words_pspy)) # sort from highest to lowest word count

# find the words per speaker per year for just the top speakers
top_words_pspy_slavery_1830 <- words_per_speaker_per_year_slavery_1830 %>%
  filter(speaker %in% unique(top_slavery_speakers$speaker)) # retain only relevant rows
```

```{r, fig.cap="Number of words spoken about slavery by individual members of Parliament from 1830 to 1838. Each point represents a speaker in a given year, plotted by the total number of words they contributed on the topic.", warning=FALSE, message=FALSE}
# visualize who spoke about slavery in 1830, how much they said, and when they spoke.
# each point represents a speaker's total words for a given year
# the x-axis shows years (rounded to whole numbers), and the y-axis shows word counts.
ggplot(top_words_pspy_slavery_1830, # make a graph
       aes(x = year, # x axis = year
           y = words_pspy, # y axis = word count
           label = speaker)) + # use speaker names for labels
  geom_text_repel(size = 2) + # repel overlapping text labels
  geom_point() + # plot each point
  ggtitle("Who spoke about slavery, how much, and when?") + # plot title
  ylab("Number of words spoken") + # y axis label
  scale_x_continuous(breaks = # whole number x axis
                       function(x) 
                         unique(floor(pretty(seq(min(x), 
                                                 (max(x) + 1) * 1))))) 
```

You may notice that we have produced a different sort of visualization
than the bar charts of the previous chapter. The ggplot library of
graphics is a highly flexible language for producing many visualizations
from the same data. Rather than the command for a bar chart --
`geom_col()` -- we have used a command for a dot plot -- `geom_point()`.
With this plot we also introduce another library, `ggrepel`, which gives
us tools for enhancing the readability of our plot by making sure that
labels do not overlap.

What does this chart of the top speakers about slavery tell us? The
chart gives us some suggestions about where to start looking. It tells
us that the five speakers who contributed the most to the debate in 1830
spoke twice as much as the speakers in other debates of the decade. The
list of these speakers gives us a good place to start looking at data.
The main speakers in 1833 were Edward Stanley, the Duke of Wellington,
James Silk Buckingham, Fowell Burton, and Robert Peel. We also see the
evidence of a second wave of debates, when issues of indentured
servitude came to parliament's attention in 1838. Then, Lord Brougham
and Lord Gleneig were the main contenders.

Wellington, Peel, Stanley, and Buckingham form a distinct set.
Wellington and Stanley had the most in common. Wellington was a former
and future prime minister and the hero of Waterloo. Stanley, the Earl of
Derby, a future prime minister, was currently in transition from his
role as Chief Secretary for Ireland to Secretary of War and the
Colonies. His career was associated with the abolition of slavery,
military rule and the establishment of schools for Ireland, and the
expansion of the vote to the middle class. Both came from noble
families; both were Whigs, although Wellington was three decades older.

Exploring the contributions of the top speakers and the tensions between
them is one area in which the tools of text mining excel.

## Using Our Skills to Explore the Historical Debates About the Abolition of Slavery

Understanding how to group data by speaker and year gives us the
possibility of closely studying the contribution of different speakers.
For instance, suppose that we wanted to read all of the debates relating
to Slavery Abolition Act of 1833, which outlawed slavery in many British
colonies, including the Caribbean and South Africa, although not across
much of India, where debt slavery would persist for decades, and where
the caste system would exclude many people from participation in the
economy through the twenty-first century.

There are some new commands below that are used for demonstration
purposes to show how one might navigate our newly joined dataset. Don't
feel worried if you can't read every line. You will be learning how to
aggregate data by speaker very soon.

```{r, message = FALSE, echo=FALSE}
# find the words spoken about slavery in the year 1833
slavery_debates_1833 <- slavery_debates_1830 %>%
  filter(year == "1833") # filter for just the year 1833 

# count how many times each speaker says each word in the slavery debates of 1833
words_per_speaker_slavery_1833 <- slavery_debates_1833 %>%
  unnest_tokens(word, text) %>% # tokenize the text column and create a new column called word
  group_by(speaker) %>% # group the data by speaker
  summarize(words_per_speaker = n()) %>% # create a new column for how many words each speaker says
  arrange(desc(words_per_speaker)) %>% # arrange in descending order
  slice(1:20) # retain just the top 20 words per speaker

# inspect the first few rows of the data to make sure it looks right
head(words_per_speaker_slavery_1833)
```

Counting how many words were spoken in the 6,000 some sentences about
the abolition of British slavery is our first reminder of the bias of
this data set: the speakers are white men, many of them of a class
invested in slavery. This collection of words is not suitable for
investigating narratives that depict lived experiences of
dehumanization, recorded from the words of enslaved people.

Nevertheless, distant reading allows us to investigate the dynamics of
politics in the British parliamentary system that made the official
abolition of slavery possible. We can take a snapshot at the speeches of
the most active speakers -- Buckingham, Stanley, Wellington, and Peel --
by comparing their most-invoked words in these debates.

In the following code, we provide a list of speakers and use the `%in%`
operator to filter for any of these speakers.

```{r, message = F}
library(knitr)
library(kableExtra)

# make a list of the names of the top speakers
pattern1 = c("Mr. Fowell Buxton", "Mr. Buckingham", "Mr. Stanley", 
             "The Duke of Wellington", "Sir Robert Peel")

# search for the words said by the speakers in pattern1 
top_slavery_speakers_1833 <- slavery_debates_1833 %>%
  filter(speaker %in% pattern1) # keep rows containing a speaker's name

top_slavery_speakers_1833 %>%
  head() %>%
  mutate(text = str_trunc(text, 120)) %>% 
  kable(format = "latex", 
        booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(4, width = "4cm") 
```

Next, let's break the text into words and clean it up, removing numbers
and stop words. We're lemmatizing each word into its word stem so that
we can count singular and plural of the same word as one.

```{r, message = F}
library(textstem)

# clean up the list of words
filtered_words_per_top_slavery_speaker_1833 <- top_slavery_speakers_1833 %>%
  unnest_tokens(word, text) %>% # break up into words 
  filter(!str_detect(word,"^\\s*[0-9]*\\s*$")) %>% # remove all numbers 
  anti_join(stop_words, by = "word") %>% # remove stopwords 
  mutate(word = lemmatize_words(word)) # lemmatize the word 

# inspect the data
head(filtered_words_per_top_slavery_speaker_1833)
```

Next, let's count the words by how many times each speaker says them.

For our visualization, we will look at only the words spoken most
frequently.

```{r, message = F}
top_words_per_abolition_speaker <- filtered_words_per_top_slavery_speaker_1833 %>%
  group_by(speaker, word) %>% # group the data by speaker 
  summarize(n = n()) %>% # count how many times each word is spoken by each speaker
  arrange(desc(n)) %>% # arrange in descending order
  slice(1:15)  # find the 15 words that each speaker says the most frequently 
  
kable(top_words_per_abolition_speaker)
```

Now visualize:

```{r, fig.cap="Top words used by six major speakers in the 1833 debates on the abolition of slavery.", echo=FALSE}
# create a plot of each speakers' favorite words 
top_words_per_abolition_speaker %>% 
  mutate(word = reorder_within(word, n, speaker)) %>% # within the grouping by speaker, rearrange the words in order by n
  ggplot(aes(x = word, # word is on the x axis
             y = n)) + # n (count) is on the y axis
  geom_col() + # make a bar chart
  facet_wrap(~speaker, # break the chart up into faceted charts, one for each speaker
             scales = "free") + # label each x and y axis separately, so that we only see the words used by that speaker on his axis
  coord_flip() + # switch the x and y axis labels (this is an elegant shortcut for other work moving labels that we'd have to do otherwise)
  scale_x_reordered() + # reorder the column before plotting with faceting 
  labs(title = "Speakers\' Favorite Words", # give a title
       subtitle = "From the 1833 Debates on the Abolition of Slavery", # give a subtitle
       x = "Word", # label the x axis 
       y = "Count") # label the y axis
```

## From Concepts to ideas

What can we learn from this visualization about the history of slavery?
How might we use the tools of text mining to learn something new about
this important historical moment?

What shows up clearly in the counts of most frequently-used words is a
series of keywords that motivated debate. The fact that all five
speakers invoke the terms "slavery," "freedom," and "labour" suggests
that they are operating as what historians call "concepts" -- the
theoretical constellations of ideas that produced much of the modern
world. Terms such as "nation," "republic," "democracy," and
"development" were argued over and redefined over the course of decades
and centuries, until the terms eventually stabilized.

Historians believe that identifying the debates over these words is key
to understanding the abstract forces that remade society, or the
"conceptual changes" by which modern understanding was produced. Peter
de Bolla and his coauthors defined concepts as "'cultural entities' held
by many agents collectively, that is held by a 'culture'." (de Bolla et.
al). They argue that concepts allow cultures to feel their way through
increasingly abstract modes of ideation, because concepts "provide the
scaffolding that enables one to understand whatever is the object of
attention and thought." In the era of the debates over the abolition of
slavery, the concept "freedom" was doing just such important work,
allowing ordinary British people and the politicians who represented
them to reckon with certain aspects of the abhorrent economic reality
that made Britain so rich in the age of empire: the fact that Britain's
wealth was largely due to the plunder of wealth from foreign lands and
especially the theft of time from the millions of enslaved humans who
harvested sugar, tea, cotton, and the other raw materials upon which
empire's trade was based. Talking about "freedom" formed the basis for
reckoning with capitalism and empire as systems, for holding them
account to a higher set of principles. That work was at its infancy in
the era of the British laws that outlawed the Atlantic slave trade and
slavery in some parts of British empire, but the work of defining and
promoting freedom continues to this day as the specific cases and
principles that comprise freedom continue to be brought into
consciousness through the work of social movements, politics, and
ideation.

Historically speaking, as concepts arise, they tend to be contested. At
the beginning, there was no single definition of "science" or "nation"
to which all parties subscribed; these terms had to be debated and
argued over until they became abstract ideas that could be invoked in
support of other debates. Corpus linguistics gives us useful clues about
how to identify potential concepts: as a rule of thumb, the words spoken
about the most in any debates are those that are most contested. In the
debates of 1833, we can tell that the words "slavery," "freedom," and
"labour" may fit such a description. Often the words spoken most
frequently are words whose definition is up for grabs; speakers who
disagree with each other will talk at great length to assert their own
definition of one of these key concepts. The two speakers who mention
freedom the most in 1833 are Buckingham and Peel, and we might expect
that they represent two poles of disagreement about what freedom is or
should be.

One part of unpacking the life of concepts is to understand the
contestation of concepts by individuals as they formulate ideas and
arguments. In this usage, we follow de Bolla and his collaborators call
the individual uses of concepts by writers and speakers "ideas" (de
Bolla et al Ch 2.1.2). Therefore, for the following discussion, let us
use "concept" to reference that shared terminology of values, especially
"freedom," while reserving "idea" for any other terms we might use to
understand how individual speakers make their cases. When people in the
past argued about how to define terms like "freedom," they used many
examples. Words such as "wages," "apprenticeship," and "regulation" were
used in the slavery debates to signal how freedom was to be achieved.

We already have enough material to hypothesize about some major
differences that marked the speakers out in their construction of the
problems of freedom and slavery. Buckingham and Buxton, representing the
abolition lobby, are more inclined to talk about the "negro" -- a term
whose usage has changed, but that in its early nineteenth-century
context, was a reference to a color -- black -- via the Latin *nigrum*.
Perhaps they were inclined to reference the individual experiences of
enslaved persons -- although this is a conjecture, something we'd need
to check in the text. Buckingham also uses words for abstract values
such as "enjoyment' and "life," where Buxton is more willing to talk
about facts on the ground in "Jamaica," for instance the "insurrection."

Problems of imperial governance concern the other speakers, whose
personal background was associated less with middle-class abolition
movements than with the military and colonial order of British empire.
Peel speaks of "colonies," "compensation," and "proprietors;" if we have
read some background to the debates, we might guess that Peel is wrapped
up with the technical questions of whether the imperial parliament in
Westminster has the right to tell the colonial government in Jamaica,
and what will be the fallout if the Westminster parliament tells the
white planters in Jamaica that their property is suddenly invalid.
Stanley's eyes are on markets -- he speaks of "sugar," the major project
produced by slave labor, "averages" (a hint that he may be speaking in
technical terms about economic production, although we would need to
read more to understand how), and "compensation" and "colonies," like
Peel. Wellington's eyes are also on imperial issues, drawing attention
(perhaps) to the implications of debates over slavery for India
(although again more reading would be necessary to understand how India
is being invoked).

Our many parenthetical remarks in the paragraph are an important
reminder that we are merely at the stage of hypothesis generation. We do
not have enough material from this list of keywords to interpret the
debates. We may begin to identify words and speakers for further
analysis.

Even in a distant reading, we can detect that Robert Peel and the Duke
of Wellington were more interested in problems of "compensation" for the
owners who benefited from the system of enslavement than in the details
of life in the West Indies -- an issue of political expediency, but also
a reflection of the ethical orientation of parliament in the 1830s.

### The Most Distinctive Words of the Top Speakers About Slavery

If we want to see the words that make each speaker distinct as opposed
to their top words, we will need a different tool than raw word count.
We might hypothesize, for instance, that Buckingham talks about
"freedom" more than the other speakers in the group, because "freedom"
is ranked as the third word for Buckingham but the eleventh for Buxton,
the fifth for Peel, and somewhere lower still for Wellington and
Stanley. But the truth is that we don't know how statistically important
these differences are without further comparison. Thinking about the
problem with greater statistical richness means asking how many times
Buckingham uttered "freedom" relative to the number of words Buckingham
spoke in the slavery debates overall; we need to know not just the full
count of one word, but how many individual words he used and how
frequently he said each, relative to the other speakers. In short,
statistics become important when we want to compare individuals.

What we need is a "distinctiveness measure," a statistical tool that
allows us to rank an index of how "distinctive" a word is of each
speaker. Further on in this book, we will go into the technicalities of
measures of distinction in greater detail, the varieties of algorithms
that can be used for this purpose, and how they work. For now, it is
enough to know that "distinctiveness" is a useful tool for understanding
which words each speakers says more relative to the other speakers.

Distinctiveness can help us to answer the question: what were the words
that Buckingham said that other speakers did not say? What were the
words that Peel uttered that other speakers never used? If raw word
count helps us to understand which concepts speakers may be arguing
over, distinctiveness helps us to interpret which keywords those
speakers used to support their argument.

Using the function `bind_tf_idf()` we can apply an index of how
distinctive each word is of each speaker. The tf-idf function will
create an index for each term-speaker pair -- a number which, if high,
tells us that the word was very distinctive of each speaker, and if low,
tell us that many of the speakers used the same word.

The `bind_tf_idf()` function takes three arguments. The first two are
the names of the facets of the data which are to be assessed for how
distinctive they are in relationship to each other -- in this case,
`word` and `speaker`. The third argument, `n`, the count of words per
speaker.

```{r, message = F}
# calculate the measure of how distinctive each keyword is of each speaker
tfidf_per_top_slavery_speaker_1833 <- top_words_per_abolition_speaker %>% 
  bind_tf_idf(word, speaker, n) %>% 
  arrange(desc(tf_idf))

# inspect the data
kable(tfidf_per_top_slavery_speaker_1833)
```

Note that the `bind_tf_idf()` function creates three new columns --
`tf`, `idf`, and `tf_idf`. The one we really care about is the last,
`tf_idf`, an index of how distinctive each word is of each speaker. A
very high `tf_idf` marks words that are said very frequently by one
speaker and never by any other speakers. A very low tf_idf marks words
that are rarely said by anyone or said by everyone with nearly the same
frequency. Looking for the high `tf_idf` words for each speaker will
tell us which words are most unique to them -- a useful clue to the
distinctive arguments, ideas, and referents that each speaker
contributed to the debate.

Now that we have an index of how distinctive each word is per speaker,
we can use this distinctiveness tool to find which words are most
distinctive of each speaker.

```{r, message = F}
# find the words most distinctive of each speaker
most_distinctive_words_per_abolition_speaker <- tfidf_per_top_slavery_speaker_1833 %>%
  group_by(speaker) %>% # group by speaker
  arrange(desc(tf_idf)) %>% # arrange distinctive words in descending order
  slice(1:15) %>% # find the top n words, ranked by tf_idf
  ungroup() %>% # ungroup the words
  mutate(index = row_number()) %>% # Add a column for each speaker’s top-ranked word
  mutate(tf_idf = round(tf_idf, 4)) %>% # round the tf_idf score to only 4 decimal places 
  select(index, word, tf_idf, n, speaker) # drop all columns except these

# create a series of smaller datasets reflecting the top speakers
speaker1 <- most_distinctive_words_per_abolition_speaker %>% 
  filter(speaker == pattern1[1]) %>% 
  select(word, tf_idf)

speaker2 <- most_distinctive_words_per_abolition_speaker %>% 
  filter(speaker == pattern1[2]) %>% 
  select(word, tf_idf)

speaker3 <- most_distinctive_words_per_abolition_speaker %>% 
  filter(speaker == pattern1[3])%>% 
  select(word, tf_idf)

speaker4 <- most_distinctive_words_per_abolition_speaker %>% 
  filter(speaker == pattern1[4])%>% 
  select(word, tf_idf)

speaker5 <- most_distinctive_words_per_abolition_speaker %>% 
  filter(speaker == pattern1[5])%>% 
  select(word, tf_idf)
```

Next, let's create a series of small tables displaying each speaker,
their most distinctive words, and the `tf_idf` score of that
word-speaker combination.

We could create a bar graph, of course, but because `tf_idf` measures
represent an abstract concept -- distinctiveness -- visualizing this
data risks confusing some possible readers, who might mistake the bar
graph of `tf_idf` for a count of frequency. Analysts often have to make
choices about how to represent their data so as to generate the least
confusion about the meaning of any statistical measure. For abstract
indexes like `tf_idf`, generated in the course of our research, we
recommend offering simple tables along with prose that explains the
significance of the numbers.

```{r, message = F}
kable(speaker1, caption = unique(most_distinctive_words_per_abolition_speaker$speaker)[1])
```

```{r, message = F}
kable(speaker2, caption = unique(most_distinctive_words_per_abolition_speaker$speaker)[2])
```

```{r, message = F}
kable(speaker3, caption = unique(most_distinctive_words_per_abolition_speaker$speaker)[3])
```

```{r, message = F}
kable(speaker4, caption = unique(most_distinctive_words_per_abolition_speaker$speaker)[4])
```

```{r, message = F}
kable(speaker5, caption = unique(most_distinctive_words_per_abolition_speaker$speaker)[5])
```

The first thing that our measure of distinctiveness helps to clarify is
what is not distinctive. The words "freedom," "labour," and "slavery"
are not particularly distinctive of any speaker -- all five speakers use
these words a great deal, and no one uses any of the words so much more
often than the others as to merit an index of the words as distinctive.

Instead, we have a new set of words that might qualify as "ideas" in the
sense that they signify the unique contribution of each of the five
speakers on the larger debate about the meanings of the concepts
"freedom" and "slavery." Words such as "enjoyment" (for Buckingham),
"wage" (for Buxton), "imperial" (for Peel), and "petition" (favored by
Wellington) give a sense of the varying priorities of individuals.

It's worth making a list of the words that intrigue us, about which we
might want to know more. It seems, for example, that each of the
speakers cultivated their own set of examples and references -- places
whose anecdotes they used as examples for thinking through the
consequences of continued slavery, the "apprenticeship" system of
continued slavery, and immediate abolition. One route through the
research might be to stop here and examine those terms and their
references, looking up one at a time:

Buxton: Hottentots, Africans, Scotchman Stanley: Venezuela, mine Peel:
Burke, Guadeloupe Wellington: Colombia

We may make many such hypotheses in the course of reading. It is always
worth writing them down; we may always choose to return to them later.
One valid use of text mining is exactly as a source for new hypotheses
about the text and how they matter.

Before we get too excited about any of these words, however, we also
need to think about the usefulness of word count and distinctiveness
together. One of the tricky things about working with tf-idf is that
it's a measure of distinctiveness, not a measure of significance. Which
is to say that a highly-ranked word from any of these lists may be so
distinctive because it was only spoken twice, and both times by one
person. Does that make the word worthy of note? Not necessarily -- it
might just be irrelevant.

We can check the significance of our distinctiveness measure by another
kind of representation of the same data, one that brings both
distinctiveness and count into the same graphic.

```{r, fig.cap="Distinctive vocabulary across six key speakers in the 1830s abolition debates. Words shown are those that appear significantly more often in one speaker’s speeches than compared with the other speakers.", message = F}

most_distinctive_words_per_abolition_speaker <- most_distinctive_words_per_abolition_speaker %>%
  mutate(word = reorder_within(word, n, speaker)) 

# x-axis: words (flipped for readability)
# y-axis: counts of each word
# fill color: tf-idf score (converted to a factor so bars are shaded categorically, not continuously)
# geom_col(): draws the bar chart
# facet_wrap(): creates separate mini-plots by speaker
# coord_flip(): makes long word labels readable
# scale_x_reordered(): preserves word ordering within facets
# scale_fill_grey(): use grayscale instead of color
# guides(fill = "none"): hides the fill legend
# scale_y_continuous(): forces whole number tick marks
# labs(): titles, subtitles, captions, axis labels
ggplot(data = most_distinctive_words_per_abolition_speaker,
       aes(most_distinctive_words_per_abolition_speaker,
           x = word,
           y = n, 
           fill = factor(tf_idf) )) +
  geom_col() + 
  facet_wrap(~speaker, scales = "free") +
  coord_flip() +
  scale_x_reordered() + 
  scale_fill_grey() + 
  guides(fill = "none") + 
  scale_y_continuous(breaks = function(x) 
    unique(floor(pretty(seq(min(x), (max(x) + 1) * 1))))) + 
  labs(title = "Speakers\' Most Distinctive Words", 
       subtitle = "From the 1830s debates on the abolition of slavery",
       caption = "Pale shading indicates greater distinctiveness", 
       x = "Word", 
       y = "Count") 
```

As opposed to the list, our bar chart shows both tf-idf distinctiveness
(represented as paleness of the bar) with count on the x axis. This
visualization can guide us as to whether we want to investigate a
particular word because of its distinctiveness -- or whether that word
might be a poor marker of significance. For instance, most of Stanley's
distinctive words are revealed to have a wordcount of only 2, which is
too few to be significant; it simply doesn't make sense that "rum,"
"mine," and "barrel" merit investigation when Stanley spoke them only
twice in his voluminous speech-giving -- that is, unless we already had
a previous interest in the rum trade and its intersection with slavery.
Statistically distinctive these words might be; worthy of further
investigation they are not (at least not necessarily).

We can use wordcount with distinctiveness to identify a set of words
that we do indeed wish to pursue further. We can see that the words
"wage," "rear," "hottentots," and "flog" are, for Fowell Buxton, both
distinctive and numerous, as are "ultimate," "qualify," "imperial," and
"king's" For Robert Peel; nearly all of Stanley's words are distinctive,
while Mr. Buckingham's most distinctive word, "thirdly," appears to be a
mere rhetorical tick.

Nevertheless, there is room for caveats about settling for these words
in particular. From the point of view of text mining, where we look to
the number of words as significant, these results are unsatisfying.
There are too many small results, where the count is smaller than 5. If
we want to write an analysis of each speakers' referents, this might be
a beginning, but we should not rest content that we have the best words
for examining the speakers yet. Let's rerun the code, this time adding
instructions to only show us words that appear above a certain threshold
by adding just a single line of code. You can play with the number in
the filter to see how the results change as we look for more and less
distinctive words.

```{r, message = F}
most_distinctive_words_per_abolition_speaker2 <- tfidf_per_top_slavery_speaker_1833 %>%
  group_by(speaker) %>%
  filter(n > 5) %>% # <----------------  this is the only thing that's new 
  arrange(desc(tf_idf)) %>%
  mutate(rank_tf_idf = row_number()) %>%
  slice(1:15) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, speaker))

most_distinctive_words_per_abolition_speaker
```

```{r, fig.cap="Most distinctive words used by six major speakers in the 1830s debates on the abolition of slavery, filtered to include only terms appearing more than five times. Pale bars indicate words that are more distinctive to each speaker relative to each other speaker.", message = F}

ggplot(data = most_distinctive_words_per_abolition_speaker2, 
       aes(x = word, y = n, fill = factor(tf_idf) )) +
  geom_col() +
  scale_fill_grey() +
  facet_wrap(~speaker, scales = "free") +
  coord_flip() +
  scale_x_reordered() + 
  guides(fill = "none") +
   scale_y_continuous(breaks = function(x) 
     unique(floor(pretty(seq(min(x), (max(x) + 1) * 1))))) + # use whole numbers on x-axis
  labs(title = "Speakers\' Most Distinctive Words with N > 5",
       subtitle = "From the Debates in the 1830s on the Abolition of Slavery",
       caption = "Pale means more distinctive",
       x = "Word",
       y = "Count")
```

With this new view, it is less clear that Mr. Buckingham wanted for
distinctive words. His use of "enjoyment," "wage," "reward," and "grow"
are notable.

At this point, we may want to choose one speaker to inspect in greater
detail. Our choice of which speaker to investigate might be motivated by
distant reading or not. We are not compelled to investigate Edward
Stanley because he spoke more than anyone else. We have used wordcount
to guide me to five possible candidates. Nor are we forced to choose
which candidate from these five to investigate further on the basis of
which words they spoke or which words the computer finds distinctive. We
can also make decisions based on outside reading, and good historians
usually do.

Suppose we choose to investigate Mr. Buckingham. He is, after all, the
outlier -- a new arrival in parliament, elected by the evangelical
middle class who obtained a presence in parliament only after 1832. We
can use this information to investigate certain aspects of his speech.
We might decide to grow curious about what he had to say about
"enjoyment," "wage," "bonage," and "america," to take a smattering of
his most distinctive words for this view.

What might we learn if we looked at Mr. Buckingham's words in greater
detail? The visualizations above limit us to a small number of words per
speaker; but we might learn a great deal more by looking at Mr.
Buckingham's favorite words, whether by count of by tf-idf. In fact, the
computation for this problem is cheap, so let's do both.

```{r, warning=FALSE, message=FALSE}
mr_buckingham_slavery_debates <- tfidf_per_top_slavery_speaker_1833 %>%
  filter(speaker == "Mr. Buckingham") %>%
  arrange(desc(n)) %>%
  slice(1:40)
  
mr_buckingham_slavery_debates
```

```{r, fig.cap="The most frequently used words in Mr. Buckingham’s speeches during the 1833 debates on the abolition of slavery. Bar shading reflects each word’s TF-IDF score, indicating how distinctive the term is relative to other abolitionist speakers. Dark coloring means the term is less distinctive, whereas the light coloring means the term is more distinctive."}

ggplot(data = mr_buckingham_slavery_debates,
  aes(x = reorder(word, n),
      y = n,
      fill = factor(tf_idf))) +
  geom_col() +
  scale_fill_grey() +
  scale_y_continuous(
    breaks = function(x) unique(
      floor(pretty(seq(min(x), max(x)))))) +
  coord_flip() +
  labs(title = "Mr Buckingham's Most Frequently Spoken Words",
       subtitle = "From the 1833 Debates on the Abolition of Slavery",
       x = "Word",
       y = "Count")
```

It is interesting that "labour" and "freedom" are used alongside
"slavery" and "slave" as Buckingham"s favorite words. Also pronounced
are his invocation of "rights," "wages," and "people," suggesting that
Buckingham was interested in drawing equivalencies between the rights of
working men in Britain and the rights of enslaved people in the West
Indies.

But none of these words are particularly distinctive of Mr. Buckingham,
at least in terms of their statistical profile. It's not clear that he
used "labour" and "freedom" any more than other speakers. In fact, his
top 50 words have a distinctiveness score of zero.

If we want to see the words that Buckingham used more than other
speakers, we can return to the tf-idf scores that we generated a moment
ago. In fact, all we have to do is change a single line of code to
produce a visualization showing the top forty most distinctive words and
their counts.

```{r, fig.cap="Mr. Buckingham’s most distinctive words in the 1833 debate on the abolition of slavery, showing the terms that appear disproportionately often in his speeches compared to the rest of the abolitionist speakers.", warning=FALSE, message=FALSE}

mr_buckingham_slavery_debates2 <- tfidf_per_top_slavery_speaker_1833 %>%
  filter(speaker == "Mr. Buckingham") %>%
  arrange(desc(tf_idf)) %>% # <------- the only line that changed
  slice(1:40)

ggplot(data = mr_buckingham_slavery_debates2,
       aes(x = reorder(word, n), # the x axis will show words ordered by their count 
           y = n),
          fill = factor(tf_idf)) + # the y axis will show n 
  scale_fill_grey() +
  scale_y_continuous(breaks = function(x) 
    unique(floor(pretty(seq(min(x), 
                            (max(x) + 1) * 1))))) + # use whole numbers for the x-axis
  geom_col() + # create a bar graph 
  coord_flip() + # switches the x and y axis for a more elegant presentation 
  labs(title = "Mr Buckingham\'s Most Distinctive Words",
       subtitle = "From the 1833 Debate on the Abolition of Slavery",
       x = "Word",
       y = "Count")
```

"Wages" emerges as a more distinctive word of Buckingham's that he said
quite frequently -- 12 times in all. It is not his most distinctive
word; to find that, we can very slightly alter the code above to find
the word he uses more than anyone else -- but the result may be
disappointing.

```{r, warning=FALSE, message=FALSE}
mr_buckingham_most_distinctive_word <- tfidf_per_top_slavery_speaker_1833 %>%
  filter(speaker == "Mr. Buckingham") %>%
  arrange(desc(tf_idf)) %>% 
  slice(1) # <------- the only line that changed

mr_buckingham_most_distinctive_word
```

"Thirdly" is merely a rhetorical habit.

More interesting are substantive words that (perhaps) indicate
Buckingham's values: wages enjoyment love universal independent devoted

-- as well as the vices he (perhaps) abhors: servitude vindictiveness
possessions ignorance grind[ing] destructive groundless cruel burthen

The "perhaps" is crucial. We are noting possible avenues for
investigation for ourselves; this is not yet a proven list of findings
from digital history. We should not be too quick to leap to conclusions
about how Buckingham used these words without reading directly from his
speeches.

```{r, message = F}
library(kableExtra)

# finding just the sentences where Buckingham uses the word "wages"
buckingham_sentences <- slavery_debates_1833 %>%
  filter(speaker == "Mr. Buckingham",
         str_detect(text, "wages"))

buckingham_sentences %>%
  head() %>%
  mutate(text = str_trunc(text, 120)) %>%   # optional: shorten long lines
  kable(format = "latex", 
        booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(4, width = "4cm")    # wrap the "text" column
```

Here they are as plaintext:

```{r}
buckingham_sentences %>%
  head() %>%
  mutate(text = str_wrap(text, 80)) %>%
  select(text) %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        longtable = TRUE, 
        escape = FALSE) %>%
  column_spec(1, width = "14cm")
```

We see in these sentences that Buckingham is using "wages" in concert
with the concepts of "freedom," "labour," and "property" to recapitulate
a Lockean argument about how property depends on the work of men to
enclose land, and how the duty fo the state is to uphold contracts that
will protect exchanges built upon this system, upon which capitalism
rests. A string of sentences follow about the specificities of wages.
But the details are hard to interpret, in part because we are missing
important context.

But before continuing further, we should observe that the sentences in
our dataset are not necessarily from a single speech or even debate. If
we look at their sentence_id numbers, we will see that the numbers are
not contiguous. These sentences came in the train of other sentences
which are missing in this dataset. To interpret any individual sentence
correctly, a data analyst must be aware of the importance of
understanding that what we are seeing here is a series of discrete
statements made on different days.

```{r, message = F}
buckingham_sentences$sentence_id
```

Sentences in Hansard are literally numbered by the order in which they
appear in print. Some of the sentences numbers above occurred near to
each other and may be from the same speech, but no sentences are
actually back-to-back in their original context. There is information
missing.

Missing information is the result of our strategy of identifying words
like "slavery" and "wages" that appear within the same sentence. For the
purpose of picking up on general patterns, this manner of counting and
generalizing is useful. But if we abstract sentences from their context,
it is almost impossible to tell what is going on.

Fortunately, we are not constrained by filtering for words. We can
always go back to the original context of the speech or debate in which
a sentence appears. We can use our list of `sentence_id` numbers to call
up the entire speeches of which each sentence is a part (further down we
will explain more about the technicalities of what happens in this
section).

If we were working on a paper analyzing these speeches, we would likely
want to read Buckingham's speeches about slavery in their entirety; the
same would be true if our research project concerned the corporate
reports of a company or a series of lawsuits. The process of reading and
understanding cannot be replaced by counting words and analyzing them.
Students of data science who pursue wordcount without background reading
inevitably arrive at weak, useless or nonsensical results.

For the moment, let us proceed by reading a limited sample of
Buckingham's speeches about slavery -- just those speeches in which he
invoked the word "wages."

The code for obtaining the speeches is given below; the full speech is
printed in the Appendix.

```{r, message = F}
# moving from a list of sentences to the full speeches for context
buck_speech_ids <- buckingham_sentences %>% 
  select(sentence_id) 

buck_speeches <- buck_speech_ids %>%
  left_join(hansard_1830, by = "sentence_id") 

buck_speeches %>%
  head() %>%
  mutate(text = str_trunc(text, 120)) %>%
  kable(format = "latex", 
        booktabs = TRUE,
        col.names=NULL,
        caption="Speeches by Buckingham about Slavery that invoke the word 'wages'") %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(2, width = "8cm")
```

With missing sentences added back in, the speeches in their entirety are
much easier to understand.

Mr. Buckingham began his long speech by introducing the three
constituencies who demanded the abolition of slavery. Some are motivated
by religion, others by claims about freedom, justice, and the British
Constitution; and still others on moral or economic grounds, which
Buckingham summarized as "policy alone."

Reviewing the call from religious circles for immediate abolition, and
from policy and constitutional circles for delay, we see Buckingham
invoking "wages" in the midst of a detailed case for immediate
abolition. He challenges the proposals in the 1834 legislation to
abolish slavery that called for a period of "apprenticeship," when
enslaved people would be nominally freed but still forced to work
without pay. The bill that finally passed would require apprenticeship,
and historians have since argued that this unwillingness to end slavery
represents a failure of the British state to commit to freedom. In
Buckingham's arguments, we see the politician as a man ahead of his time
-- arguing that "apprenticeship" or "servitude" was an undisguised
synonym for slavery, and denouncing any act that extended the period of
enslavement.

We also see, in the second speech, we see Buckingham extraoplating from
his experience of working with forced labor on naval vessels to the
question of apprenticeship. We see him calling up another one of his
more distinctive keywords -- "America" -- to make the case that the
independent United States provided a historical example of the success
of freeing an expensive colony to self-government.

Buckingham's commitments to endorse the humanity and rights of enslaved
humans are abundantly clear. We see him making a detailed argument
against the category of race itself as the basis for political or
economic assumptions, especially assumptions about inferiority. He makes
a case for dark-skinned people freely associating and participating in
the economies of the Mediterranean, invoking ancient Egypt and the
biblical Queen of Sheba too argue that dark-skinned Africans and their
interracial descendents were destined for economic success.

We see Buckingham arguing that the free market naturally rewards
laborers who choose dangerous work. Buckingham later generalizes about
the rewards of a free market from his experience in Sheffield observing
steel grinders and the high wages they commanded. Buckingham is arguing
that only the freedom of wages -- not apprenticeship -- will induce
formerly enslaved men to take on challenging work.

In this reading, the word "wages" is a category that allows Buckingham
to navigate between his concerns and those of other policy-makers.
Buckingham's invocation of wages suggest that he believes that
capitalism has a utopian side in the freedoms it offers to workers. His
anecdotes about Sheffield's steel workers and British naval vesslels
suggest that he...

But we are not done. We can use the distinctive words in our list to
investigate other aspects of the speech.

```{r, message = F}
buckingham_sentences_2 <- slavery_debates_1830 %>%
  left_join(speaker_metadata_1830) %>%
  filter(speaker == "Mr. Buckingham",
         str_detect(text, "enjoyment"))
```

```{r}
buckingham_sentences_2 %>%
  head() %>%
  mutate(text = str_wrap(text, 80)) %>%
  select(text) %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        longtable = TRUE, 
        escape = FALSE) %>%
  column_spec(1, width = "14cm")
```

If the example of "wages" furnished sentences difficult to understand
without broader context, the sentences about "enjoyment" give an example
where Buckingham's case is clear without any further effort peering into
the text. His utopian vision of capitalism becomes clear in these
passages, which posit the conjecture that the freedom to work on a wage
market gives labourers access to "enjoyment," and that enjoyment and its
natural incentives therefore offers a reason why the slaves should be
freed.

```{r, message = F}
buckingham_sentences_3 <- slavery_debates_1830 %>%
  left_join(speaker_metadata_1830) %>%
  filter(speaker == "Mr. Buckingham",
         str_detect(text, "love"))
```

```{r}
buckingham_sentences_3 %>%
  head() %>%
  mutate(text = str_wrap(text, 80)) %>%
  select(text) %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        longtable = TRUE, 
        escape = FALSE) %>%
  column_spec(1, width = "14cm")
```

We might also elect to investigate "love," one of the words that
appeared highly ranked through tf-idf but relatively sparse. We
investigated the word because we felt that it might represent one of a
set of virtues to which Buckingham looked -- alongside "universal,"
"desire," and "religion."

This is a dead end. When we consult the text, we do not find an
endorsement of "love" as a universal virtue that we hypothesized when we
looked at the word list. Instead, we find the use of phrases such as
"love of ease," "love of liberty," and "love of repose," as Buckingham
invokes conjectures about the motivations of humankind, endorsing none
of them. It would be foolhardy to make any kind of a positive argument
about statements so diverse as these. We could try to unify them under
one statement, but we would be falsifying evidence. In a research paper
we will have no use for these examples, and we will pass over the word
"love" without remark.

```{r, message = F}
buckingham_sentences_4 <- slavery_debates_1830 %>%
  left_join(speaker_metadata_1830) %>%
  filter(speaker == "Mr. Buckingham",
         str_detect(text, "universal"))
```

```{r}
buckingham_sentences_4 %>%
  head() %>%
  mutate(text = str_wrap(text, 80)) %>%
  select(text) %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        longtable = TRUE, 
        escape = FALSE) %>%
  column_spec(1, width = "14cm")
```

We find in these sentences Buckingham invoking "universal" in the sense
of an Enlightenment mandate, whether an expression of consensus or
natural laws. Democratic demand for the abolition of slavery is an
expression of a "universal" sentiment (sentences 1, 3 and 4).
"Universal" laws govern behavior, including the fact that rebellion is
more scarce in free nations (sentence 2). Universal truths, revealed
through democratic processes, give the basis for legislation, according
to Buckingham.

Presented with these particular sentences, having read some of
Buckingham's arguments at greater length, we may feel confident that we
understand how he has used the words. As we are making a wider argument
about the use of words, buttressed by several examples, we do not
necessarily need to follow every single sentence back to its original
context.

```{r, message = F}
buckingham_sentences_5 <- slavery_debates_1830 %>%
  left_join(speaker_metadata_1830) %>%
  filter(speaker == "Mr. Buckingham",
         str_detect(text, "desire"))
```

```{r}
buckingham_sentences_5 %>%
  head() %>%
  mutate(text = str_wrap(text, 80)) %>%
  select(text) %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        longtable = TRUE, 
        escape = FALSE) %>%
  column_spec(1, width = "14cm")
```

Buckingham invokes "desire" as the basis for understanding. Buckingham
invokes universal desire in the spirit of Locke, Mandeville, and Smith
as forces that incline men to leisure and pleasure and thus incentivize
work (sentences 1, 2, 3, 6, 7). Buckingham suggests that former enslaved
persons could eventually follow the American model, growing wealthy
enough to become consumers who would buy finished products from British
manufactures. A process of mutual enrichment would be the consequence
(sentence 4).

The sole sentence we can't understand without broader context is
Sentence #5. Let's inspect the sentence to obtain the sentence ID
number.

```{r}
buckingham_sentences_5[5]
```

The id of Sentence #5 is S3V0018P0_8993. Since we understand the gist of
Buckingham's arguments about slavery, we might first try looking at the
sentences immediately preceding this passage. The sentences are numbered
chronologically, so all we have to do is to pull a few sentence numbers
starting with S3V0018P0_8990.

```{r}
missing_context <- hansard_1830 %>% 
  filter(sentence_id %in% c("S3V0018P0_8990", "S3V0018P0_8991", "S3V0018P0_8992", 
                            "S3V0018P0_8993"))

missing_context %>%
  head() %>%
  mutate(text = str_wrap(text, 80)) %>%
  select(text) %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        longtable = TRUE, 
        escape = FALSE) %>%
  column_spec(1, width = "14cm")
```

In Sentence #5, Buckingham is talking about how workers will pursue
dangerous and difficult jobs when motivated by desire ("gain, honour,
promotion, and prize-money.") For this reason, Buckingham concluded that
"promotion was the desire of all," and that, if allowed to pursue them,
the prospect of wages would incentivize formerly enslaved people as it
had incentivized British navies to work hard in the pursuit of earnings.

### Pulling Together an Analysis

In the foregoing exercises, we have investigated many dimensions of text
use in the debates over slavery. We have come up with hypotheses and
tested some (if not all) of them. We have read words in context,
returning to the text for more sentences when we needed more
information. The material that we now have is enough to make an argument
about Buckingham's use of words to make an argument about capitalism
that supported the immediate abolition of slavery.

We have applied our personal judgment in deciding when word lists are
enough by themselves, when a few sentences can inform our understanding,
and when we need to return to the speeches themselves for more context.
Cultivating the judgment to know when to trust data, when to look for
more context, and when to leave out an example entirely is one of the
most important skills of the digital historian -- and it is a skill
almost never tught in departments of Comptuer Science or Data Science.
Understanding which course to pursue is a matter of training, instinct,
broad reading, and a willingness to engage all the tools. The
interpreter who uses word count skilfully has no automatic answer for
how to interpret a given speaker or given text; the top words are no
sure guide, nor is tf-idf, nor the sentences linked to them. Only by
knowing when to move from short sentences to longer speeches, reading
the speeches in context, and examining the words thoughtfully is an
adequate opinion of the text formed.

How much of the work should I show in this case? The answer may be
surprising. None of them.

The process of testing hypothesess laid out here is considered part of
the research. Most of the visualizations and discussion would not form
part of a term paper, let alone a published paper. They are like the
wood shavings that fill a workshop; they are evidence of the labor that
has gone into making an ultimate artifact at which we aim, which is
surprising knowledge about the past. At the end of the day, the shavings
are swept from the workshop floor to be discarded or composted. They are
not to be treasured or preserved. Just so with much of the intermediate
work of reading, counting, and modeling that goes into the making of an
insightful historical analysis. There is no shortcut through it. No
language model gives us a substitute for thinking about words and their
meaning.

The graphs we made heretofore give us context that would otherwise be
invisible. If not for text mining, we would have little idea that
Buckingham and Buxton both spoke about wages disproportionately to
Stanley and Peel. We might select other terms through close reading that
catch our eye, and we might write about how the speakers used these
words to support an argument. In any case, that is the task ahead of us
— to show how speakers used their words to make a political case. But
without text mining, the basis for the selection of words would be
hopelessly colored by our individual prejudices. We can make a
fundamentally more objective case for which words we focus on when we
approach the problem through text mining.

To make an argument about Buckingham's use of words, what we need is (1)
a list of the words we chose to investigate and a statement of the role
that text mining played in selecting them, and (2) a close reading of
the words in context that proves that we read Buckingham's speeches.

For the purpose, we do not need to show the reader the output of our
tf-idf or word count process. Instead, a list of words will suffice. In
the prose, we would spend a sentence or two describing the process we
went through of applying tf-idf with and without a threshold to inspect
Buckingham's distinctive words when compared with four other top
speakers in the debates. If we want the list formatted nicely, we might
write some code like this:

```{r}
# concatenate a list of the words I'm interested in
my_words <- c("wage", "enjoyment", "love", "universal", "desire")

# find the tfidf scores for my words by looking in previous data
my_words_w_tfidf <- tfidf_per_top_slavery_speaker_1833 %>% 
  filter(speaker == "Mr. Buckingham", word %in% my_words) %>% # filter for words by Buckingham
  ungroup() %>% # necessary because the previous dataset was grouped
  select(word, n, tf_idf) %>% # retain three columns
  mutate(tf_idf = round(tf_idf, 4)) # round tf-idf to four decimal places

# make a table
kable(my_words_w_tfidf,
      format = "latex",
      booktabs = TRUE,
      caption = "Some of Mr. Buckingham's Most Distinctive Words in Debates About Slavery")
```

As to the substance of the argument, we have most of the interpretive
work above in the sentences that move, word by word, through a synopsis
of the arguments Buckingham made about slavery, through his use of the
words "desire" and "universal" to sketch out the enlightenment system of
perfection that he believed made immediate abolition both inevitable and
desirable, and finally through his identification of "wages" as the
crucial mechanism that made it possible for men's willingness to endure
dangerous work to be rewarded and thus workers to be incentivized to
learn and apply themselves -- something slavery could not provide. We
might cut and paste into a new series of paragraphs our piecemeal
analyses above of words in context. We might reshape the resulting
paragraph with topic and concluding sentences into a series of arguments
supported by evidence. The resulting argument, based on the actual words
of the primary text, will be -- above all -- highly persuasive, because
it is composed solely of arguments and pieces of evidence, each argument
corresponding to a series of pieces of evidence.

At the risk of repetition, let us review this key point: text mining is
not meant to save us from reading. It is meant to help us read better.
The positive result of text mining is a precise understanding of the
words used to make particular arguments by historical actors, alongside
a statistical understanding of which words mark out individuals from
their peers. The best outcome of engaging in a text mining project
should be judged not by the fanciness of the graph that we produce but
by the seriousness with which the analyst engages original texts and
their language.

To pursue a question seriously means being willing to walk into blind
alleys in search of some historical surprise. The frustrating fact of
our work so far is that it has not produced anything astonishing. That
advocates of the abolition of slavery invoked what they believed to be
the utopian aspects of capitalism as an argument has been noted by
previous generations of historians. That Buckingham had a specific
vocabulary with which to engage capitalism -- enjoyment, wages, the
universal, and desire being its keywords -- may be a new finding, but
it's hardly headline news.

The work we have performed thus far gives us the material for a 3-page
paper that demonstrates competence but not insight.

To move forward towards insight, we would want to return to our
hypotheses and perform more comparisons. Whether the goal is a five- or
ten-page paper or a corporate report, we will need to demonstrate that
our method can discover something fundamentally new.

### Next Steps: From Competence to Insight

Fortunately, we have several hypotheses ready to hand. We could begin by
testing Buckingham's use of the word "wages" against Buxton's use of the
same term -- which Buxton used an astonishing 40 times in the course of
the slavery debates. Did they speak about wages in the same way? Did
they make the same argument? But if our findings about wages challenge
little that historians already know about utopian faith in capitalism at
the beginning o the nineteenth century, this looks like an unpromising
direction.

Alternatively, we might entertain our hypotheses about the other key
speakers. We might analyze the distinctive words of the other speakers
-- Stanley, Peel and Wellington, whose speeches offer such a different
perspective on the debates. For each subject being analyzed, we would
use the same code, making minor adjustments to allow us to pull the
sentences of each speaker and inspect them in greater detail.

One place to start is Peel's invocation of the word "assembly," a word
that in this context refers to the local government of the West Indies,
including Jamaica. These island assemblies were unique among the British
colonies in having the kind of independence that the North American
colonies desired but were denied. They had an independent legislature of
their own, which had operated with relative independence since the
seventeenth century. One question raised by the abolition of slavery was
whether these assemblies would continue to legislate in the West Indies,
and whether it might become possible for formerly enslaved people to
have a vote or stand for representation in them -- that is, whether it
might be possible for persons of color to control their own political
destiny.

Most of the speakers didn't use the word "assembly" or used it rarely.
But for Peel, "assembly" was a distinctive word, used 11 times in 1833
alone.

```{r, message = F}
# merge slavery_1830 with speaker metadata
# keep only sentences by Peel
# str_detect to search for capitalized, uncapitalized, singular and plural
peel_sentences <- slavery_1830 %>%
  left_join(speaker_metadata_1830) %>%
  filter(speaker == "Sir Robert Peel",
         str_detect(text, "Assembl|assembl"))
```

```{r}
peel_sentences %>%
  head() %>%
  mutate(text = str_wrap(text, 80)) %>%
  select(text) %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        longtable = TRUE, 
        escape = FALSE) %>%
  column_spec(1, width = "14cm")
```

Here again we must read the sentences in their original speeches to make
sense of it.

```{r, message=FALSE}
# moving from a list of sentences to the full speeches for context
# Trim each source to just the fields you need
peel_base <- peel_sentences %>%
  select(sentence_id)

hansard_trim <- hansard_1830 %>%
  select(sentence_id, text)

debate_meta_trim <- debate_metadata_1830 %>%
  select(sentence_id, speechdate)

speaker_meta_trim <- speaker_metadata_1830 %>%
  select(sentence_id, speaker)

file_meta_trim <- file_metadata_1830 %>%
  select(sentence_id, speech_id)
```

```{r}
peel_speeches <- peel_base %>%
  left_join(hansard_trim,  by = "sentence_id") %>%
  left_join(speaker_meta_trim,  by = "sentence_id") %>%
  left_join(file_meta_trim, by = "sentence_id") %>% 
  left_join(debate_meta_trim, by = "sentence_id") %>%
  group_by(speech_id, speaker, speechdate) %>%
  summarise(speech = paste(text, collapse = " "), .groups = "drop") %>%
  mutate(speech = paste0(speechdate, ": ", toupper(speaker), " ", speech)) %>%
  select(speech)

peel_speeches %>%
  head() %>%
  mutate(speech = str_trunc(speech, 120)) %>% 
  kable(format = "latex", 
        booktabs = TRUE, 
        caption = "Three Speeches by Buckingham about Slavery Invoking the Word 'Wages'") %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(1, width = "14cm") 
```

In the three speeches above, we find Peel invoking the colonial
assemblies of the West Indies in order to defend the principle of
independent democracies being entitled to make their own laws,
unhindered by empire -- even when the laws contained principles so
odious as that of the enslavement of human beings.

In the first speech, April 15, 1833, Peel noted his dissatisfaction with
the Colonial Assemblies but urged moderation and warned against making
hasty resolutions that could harm the interests of slaves, humanity, and
planters.

In the second speech, May 24, 1833, Peel pointed out the danger of
making hasty regulations from afar without understanding the local
conditions in the West Indies, especially in light of recent
insurrections in Jamaica. He warned against inciting slaves to
resistance through "incautious language and false hopes," only to then
suppress them with military force, which would be unjust and exacerbate
the situation.

Cautioning against haste, Peel invoked the assemblies, suggesting that
it was a desire to "triumph over" the planters in their assemblies that
made advocates of abolition impatient. Against haste, he advised a slow
and deliberate plan to ensure order -- the plan for apprenticeships,
which would continue slavery under another name for seven further years.

In the third speech, June 3, 1833, Peel again urged his listeners not to
interfere in the affairs of colonial assemblies, "but to lay the
foundation of future prosperity and tranquillity in those countries."

He raised questions about whether slaves were "fit for freedom." The
crowning point in Peel's argument was a lesson about history. Peel
shocked his listeners by alluding to the liberation of the slaves in the
French colony of St. Domingo in 1794, which had been followed riots
which Peel described as "atrocities," which resulted in the leveling of
inhabitations and the enslavement of whites by black pirates. In his
account, Peel argued that the violence was levied against white bodies
by black ones, and that the same specter of black violence awaited
Englishmen in the West Indies if they moved towards liberation too
quickly.

The existence of slavery thus became the basis for arguing that the West
Indies could have no possible future as an independent government:

> "When a state of slavery is that upon which we are to work, the very
> means which lead to liberty must partake of compulsion. The minds of
> men being crippled with that restraint, can do nothing for themselves;
> every thing must be done for them. The regulations can owe little to
> consent."

When we compare Buckingham's arguments about the "universal" with Peel's
arguments about "assembly," what we see is a standoff between two
versions of the enlightenment, both full of faith for hope in humanity
and the power of reason, but which worked against each other in a
dramatic standoff.

The first version -- Buckingham's -- defined freedom in economic terms.
It found slavery abhorrent and declared that it must be immediately
desired. Buckingham's contention that the laws of unfreedom, once
righted, would turned former slaves into productive subjects was
premised upon faith in the laws of the free market and the ordered
workings of desire, which envisioned formerly enslaved persons becoming
free and going on to pursue dangerous and rewarding work alongside the
British working class.

The second version -- Peel's -- defined freedom as the ability of
educated individuals to elect their own leaders. It defined the basis
for democracy as sacroscanct. At the same time, it held up a standard
for being fit to govern. Only educated, peaceful people could be allowed
to elect their leaders. This bar to entry -- the defining of the
conditions under which liberal individuals would be fit to govern -- was
one of the major points that would delimit democracy for most of the
nineteenth and twentieth-century experiments with democracy. Again and
again, points of this kind would be invoked to artificially limit the
freedoms of individuals (like the enslaved inhabitants of the West
Indies). And the argument was composed of a slight-of-hand: on the one
hand, the endorsement of democracy (and the colonial assembles) as
sacrosanct places of law-making whose powers must not be challenged by
an imperial authority. On the other hand, the fear of unruly
individuals, driven by race-based terror, backed by anecdotes about
black pirates. The upshot of this argument was the denial that formerly
enslaved persons could be educated for granted independence, who looked
to the coercive power of parliament alone for saftey.

Both Buckingham's and Peel's arguments about freedom were premised on
rationality, humanity, and the idea that the duty of parliament was to
promote peace around the world. But it was the latter -- the argument
that democracy could only succeed where coercion was applied to keep
unruly forces at bay -- that won the day.

Here, at last, an argument wider than mere personality, rhetoric, or
affectation is starting to emerge. It is likely that if we pursued
further looking into the top speakers and their favored words, we would
be able to identify still further complcations and aspects of the
enlightenment's encounter with slavery in 1833. The process would be the
same as it was above: using code to move from words to sentences and
sentences to their context, then reading, summarizing, and interpreting
how words were shaped by speakers into arguments and positions that
encapsulate the differences of world view at work in 1833 and the
terrible consequences that flowed from the failures of that moment.

What makes this version of the argument "insightful". in comparison to
the others? Put simply, an argument should reflect back on all possible
levels of analysis -- from the individual in their context to the
global, political, and economic dimensions of the historical debate. We
turn to historical analysis, as Quentin Skinner puts it, to act "as a
kind of archaeologist, bringing buried intellectual treasure back to the
surface, dusting it down and enabling us to consider what we think of
it." (Skinner, Liberty Before Liberalism (1998), p. 112). In order to
have an informed opinion about the positions of Buckingham and Peel
around slavery -- and therefore in order to assess the ways in which the
liberties of enslaved persons were postponed again and again -- it is
useful to have statistical insights into the keywords "universal" and
"assemblies" upon which Buckingham and Peel staked their claims about
freedom. We have not done the work when we arrive at lists of keywords,
however. The keywords must be read and understood in context. The
arguments must be summarized. Their implications must be considered.
Above all, we must strive to grasp at an explanation that helps us to
understand why the abolition of slavery was so politically difficult for
the seemingly-well intentioned men who gathered in Westminster in 1833
-- and what we can learn about reasoned claims about government from
that date. If the keywords do not unlock some insights about that
moment, they do us no good.

The claim of this book has been that the application of historical
reasoning with text mining is useful not only for professional studies
of history but also for the many applied areas where understanding the
reaction to some crisis is useful -- a theme imprtant in business
management, journalism, and politics. The crisis we wish to interpret
when text mining corporate reports or journalistic stories might not be
the postponed abolition of slavery in British Empire. But it will still
be useful for us to understand the significance of individual points of
view -- for instance, memos written by separate executives or
originating in different wings of the government -- and how they
instrumentalize particular keywords as anchors for broader concepts, be
those concepts "freedom" and "slavery" or "earnings" and "employment."

### Did the Concept of "Freedom" Change After the Debates of 1833?

The reason why the top speakers in any debate matter is the hypothesis
that their words may have changed the minds of many listeners. Spoken
aloud before an audience of other members of the press, reprinted in the
magazines and newspapers of the nation, the words spoken in parliament
mattered.

We can ask this question two ways: first, by asking about the debates
around slavery before and after 1833, and secondly by asking the same
question about all of Hansard for the same dates.

```{r, message=FALSE}
slavery_debates_before_1833 <- slavery_1830 %>% 
  filter(year < 1833) 

freedom_collocates_before_1833 <- slavery_debates_before_1833 %>%
  mutate(when = "1830-32") %>%
  filter(str_detect(tolower(text), "labour")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  group_by(when, word) %>%
  reframe(n = n()) %>%
  arrange(desc(n)) 

top_freedom_collocates_before_1833 <- freedom_collocates_before_1833 %>%
  top_n(30) 

top_freedom_collocates_before_1833
```

```{r, message=FALSE}
slavery_debates_after_1833 <- slavery_1830 %>% 
  filter(year > 1833) 

freedom_collocates_after_1833 <- slavery_debates_after_1833 %>%
  mutate(when = "1834-36") %>%
  filter(str_detect(tolower(text), "free")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  group_by(when, word) %>%
  reframe(n = n()) %>%
  arrange(desc(n)) 

top_freedom_collocates_after_1833 <- freedom_collocates_before_1833 %>%
  top_n(30) 

top_freedom_collocates_after_1833
```

Let's use tf-idf to find the words that are most distinctive of each
period -- before and after 1833 -- in the debates about slavery.

```{r, message=FALSE}
freedom_collocates_before_1833 <- slavery_debates_before_1833 %>%
  mutate(when = "1830-32") %>%
  filter(str_detect(tolower(text), "free")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>% 
  group_by(when, word) %>%
  reframe(n = n()) %>%
  arrange(desc(n)) 

freedom_collocates_after_1833 <- slavery_debates_after_1833 %>%
  mutate(when = "1834-36") %>%
  filter(str_detect(tolower(text), "free")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  group_by(when, word) %>%
  reframe(n = n()) %>%
  arrange(desc(n)) 


tf_idf_freedom_collocates <- bind_rows(freedom_collocates_after_1833, 
                                       freedom_collocates_before_1833) %>%
  bind_tf_idf(word, when, n) %>%
  arrange(-tf_idf) %>%
  select(when, word, n, tf_idf)

tf_idf_freedom_collocates_before <- tf_idf_freedom_collocates %>%
  filter(when == "1830-32") %>%
  select(-when) %>%
  top_n(15)

tf_idf_freedom_collocates_after <- tf_idf_freedom_collocates %>%
  filter(when == "1834-36") %>%
  select(-when) %>%
  top_n(15)
```

```{r}
kable(tf_idf_freedom_collocates_before, caption = "1830-32")
```

```{r}
kable(tf_idf_freedom_collocates_after, caption = "1834-36")
```

Let's look to see if freedom's meaning changed across everything.

```{r, warning=FALSE, message=FALSE}
hansard_before_1833 <- hansard_1830 %>% 
  left_join(debate_metadata_1830, by = "sentence_id") %>%
  mutate(year = year(speechdate)) %>%
  filter(year < 1833) %>%
  bind_rows(hansard_1820)

hansard_after_1833 <- hansard_1830 %>%
  left_join(debate_metadata_1830, by = "sentence_id") %>%
  mutate(year = year(speechdate)) %>%
  filter(year > 1833)

freedom_collocates_before_1833 <- hansard_before_1833 %>%
  mutate(when = "before 1833") %>%
  filter(str_detect(tolower(text), "freedom")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>% 
 group_by(when, word) %>%
  reframe(n = n()) %>%
  arrange(desc(n)) 


freedom_collocates_after_1833 <- hansard_after_1833 %>%
  mutate(when = "after 1833") %>%
  filter(str_detect(tolower(text), "freedom")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
 group_by(when, word) %>%
  reframe(n = n()) %>%
  arrange(desc(n)) 

tf_idf_freedom_collocates <- bind_rows(freedom_collocates_after_1833, 
                                       freedom_collocates_before_1833) %>%
  bind_tf_idf(word, when, n) %>%
  arrange(-tf_idf) %>%
  select(when, word, n, tf_idf)

tf_idf_freedom_collocates_before <- tf_idf_freedom_collocates %>%
  filter(when == "before 1833") %>%
  select(-when) %>%
  #filter(n>5) %>%
  top_n(15, wt = tf_idf)

tf_idf_freedom_collocates_after <- tf_idf_freedom_collocates %>%
  filter(when == "after 1833") %>%
  select(-when) %>%
  top_n(15, wt = tf_idf)
```

```{r}
kable(tf_idf_freedom_collocates_before, caption = "1820-32")
```

```{r}
kable(tf_idf_freedom_collocates_after, caption = "1834-39")
```

# EXERCISES------------------------------------------

## 1. In the section, “Did the Concept of "Freedom" Change After the Debates of 1833?,” we don’t give an interpretation of the tables showing the collocates of the word “freedom over time. Examining the results, what do you see? Did the meaning of freedom change in these years? How? Specify as many words as possible in your answer. Adjust the number of words displayed to do your research.

```{r freedom_1883, echo = FALSE}

top_freedom_collocates_after_1833 <- freedom_collocates_after_1833 %>%
  slice_max(n, n = 80, with_ties = FALSE)

top_freedom_collocates_before_1833 <- freedom_collocates_before_1833 %>%
  slice_max(n, n = 80, with_ties = FALSE)

tf_idf_freedom_collocates_before <- tf_idf_freedom_collocates %>%
  filter(when == "before 1833") %>%
  select(-when) %>%
  slice_max(tf_idf, n = 40, with_ties = FALSE)

tf_idf_freedom_collocates_after <- tf_idf_freedom_collocates %>%
  filter(when == "after 1833") %>%
  select(-when) %>%
  slice_max(tf_idf, n = 40, with_ties = FALSE)

kable(top_freedom_collocates_before_1833, format = "latex", booktabs = TRUE)
kable(top_freedom_collocates_after_1833, format = "latex", booktabs = TRUE)
kable(tf_idf_freedom_collocates_before, format = "latex", booktabs = TRUE)
kable(tf_idf_freedom_collocates_after, format = "latex", booktabs = TRUE)

```

Before 1833, the most common collocates appearing in the same sentences
as “freedom/free” were institutional and constitutional vocabulary -
such as “parliament”, “constitution”, “law”, “trade”, “petitioners”, and
“England” - suggesting freedom is often framed through parliamentary
procedure, petitions, and constitutional order. After 1833, the
collocates remain political (house, country, government, parliament),
but also start to shift toward a more overtly electoral and partisan
frame: “election” rises sharply, and terms like “noble,” “lord,”
“rights,” and “religious” appear among the top collocates. This analysis
shows that “freedom” remains a political concept both before and after,
but over time it shifts more toward elections, the rights of the people,
and contested public politics than toward petitions and constitutional
framing alone. At the same time, as the chapter notes, some of this change may reflect differences in the types of speeches captured in each period rather than a complete semantic shift, since "slave freedom" and "civic freedom" can also appear together in the same debates.

## 2. In this chapter, we used distinctiveness to find the top speakers about the British debates around the abolition of slavery in 1833. How would you use distinctiveness — tf-idf — to investigate the different attitudes about the next most important debate of the period, the debate over who qualified for welfare in the lead up the Poor Law Amendment of 1834?

```{r poorlaw, echo = FALSE, results='asis'}
# Join text + debate metadata
poorlaw_1830 <- hansard_1830 %>%
  left_join(debate_metadata_1830, by = "sentence_id") %>%
  mutate(
    year = year(speechdate),
    debate = iconv(debate, from = "", to = "UTF-8", sub = "")
  ) %>%
  filter(year %in% 1833:1834) %>%
  filter(str_detect(
    debate,
    regex("\\bpoor\\s+law\\b|\\bpoor\\s+rate(s)?\\b|\\bworkhouse(s)?\\b|\\bpauper(s)?\\b|\\bout-?door\\s+relief\\b",
          ignore_case = TRUE)
  ))

# Add speakers
poorlaw_debates <- poorlaw_1830 %>%
  left_join(speaker_metadata_1830, by = "sentence_id") %>%
  filter(suggested_speaker != "") %>%
  filter(str_detect(
    text,
    regex("\\b(pauper(s)?|poor\\s+rate(s)?|relief|workhouse(s)?|settlement|removal|able-?bodied)\\b",
          ignore_case = TRUE)
  )) %>%
  filter(!str_detect(
    text,
    regex("\\b(church|tithe(s)?|rector|clergyman|dean|incumbent)\\b",
          ignore_case = TRUE)
  )) %>%
  select(sentence_id, speaker, suggested_speaker, text, year, debate)

# Tokenize + clean 
poorlaw_words <- poorlaw_debates %>%
  unnest_tokens(word, text) %>%
  filter(!str_detect(word, "^\\s*[0-9]*\\s*$")) %>%
  anti_join(stop_words, by = "word")

top_poorlaw_speakers <- poorlaw_words %>%
  group_by(speaker) %>%
  summarize(words_per_speaker = n(), .groups = "drop") %>%
  arrange(desc(words_per_speaker)) %>%
  slice(1:10)

# Compute tf-idf 
poorlaw_tfidf <- poorlaw_debates %>%
  filter(speaker %in% top_poorlaw_speakers$speaker) %>%
  unnest_tokens(word, text) %>%
  filter(!str_detect(word, "^\\s*[0-9]*\\s*$")) %>%
  anti_join(stop_words, by = "word") %>%
  group_by(speaker, word) %>%
  summarize(n = n(), .groups = "drop") %>%
  filter(n >= 2) %>%   # was >= 5
  bind_tf_idf(word, speaker, n) %>%
  arrange(desc(tf_idf))

poorlaw_tfidf %>%
  select(speaker, word, n, tf_idf) %>%
  group_by(speaker) %>%
  slice_max(tf_idf, n = 25, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(tf_idf = round(tf_idf, 4)) %>%
  knitr::kable(format = "latex", booktabs = TRUE,
               caption = "Poor Law debates (1833–34) - top distinctive words by speaker (tf-idf)") %>%
  kableExtra::kable_styling(full_width = FALSE,
                            latex_options = c("hold_position", "scale_down", "repeat_header"))
```

The tf-idf results show that MPs approached the question of who deserves
welfare using different frameworks, not just different opinions. Sir
Samuel Whalley stands out with his use of administrative and
cost-related terms like “paupers,” “removal,” “parish,” “rate,”
“relief,” “chargeable,” “expense,” “county/counties,” and “magistrates.”
This pattern suggests he viewed welfare less as a moral issue and more
as a question of local governance. He focused on deciding who is
“chargeable” to which “parish” or “county,” how “removal” should be
handled, and how to manage the “rate” and total “expense” of supporting
“paupers.” Other speakers focus on different aspects, even within the
same welfare system. For example, Mr. Barron is associated with
“Ireland,” “country,” and “poor,” while Mr. Robert Palmer stands out for
often using “irish” along with “paupers.” This suggests that welfare
eligibility and responsibility were often discussed in terms of Irish
poverty and mobility. Overall, tf-idf shows us that the debate was not
just about more relief vs less relief, but about different ways of
defining the policy problem. There were disputes over local
responsibility and legal administration (“parish,” “removal,”
“chargeable,” “magistrates”) compared to broader national or Irish
perspectives (“Ireland,” “irish,” “country,” “poor”).

As we saw in Chapter 4’s analysis of slavery speakers, the next step
would be to pull context sentences and rebuild speeches around these key
terms (such as “removal,” “chargeable,” “paupers,” and “rate”) to see
how MPs explained who should get assistance and which local units should
pay for it.

## 3. In this chapter, we investigated the speeches of Mr. Buckingham. Choose one of the other top speakers on slavery and investigate his favorite words, calling up tables to show how those words were used in context. Share the tables you create. Read the related speeches, and explain what you have learned about how Buckingham and your speaker are different.

```{r favespeakr, echo = FALSE}
peel_top_by_count <- tfidf_per_top_slavery_speaker_1833 %>%
  filter(speaker == "Sir Robert Peel") %>%
  arrange(desc(n)) %>%
  slice(1:25)

peel_top_by_tfidf <- tfidf_per_top_slavery_speaker_1833 %>%
  filter(speaker == "Sir Robert Peel") %>%
  arrange(desc(tf_idf)) %>%
  filter(n >= 5) %>%
  slice(1:25)

### peel anchor word

peel_sentences <- slavery_debates_1833 %>%
filter(speaker == "Sir Robert Peel",
       str_detect(text, regex("\\bassembl(y|ies)\\b", ignore_case = TRUE)))%>%
  select(sentence_id, speaker, text, year)

peel_base <- peel_sentences %>%
  select(sentence_id)

hansard_trim <- hansard_1830 %>%
  select(sentence_id, text)

speaker_meta_trim <- speaker_metadata_1830 %>%
  select(sentence_id, speaker)

file_meta_trim <- file_metadata_1830 %>%
  select(sentence_id, speech_id)

debate_meta_trim <- debate_metadata_1830 %>%
  select(sentence_id, speechdate)

peel_speeches <- peel_base %>%
  left_join(hansard_trim, by = "sentence_id") %>%
  left_join(speaker_meta_trim, by = "sentence_id") %>%
  left_join(file_meta_trim, by = "sentence_id") %>%
  left_join(debate_meta_trim, by = "sentence_id") %>%
  group_by(speech_id, speaker, speechdate) %>%
  summarise(speech = paste(text, collapse = " "), .groups = "drop") %>%
  mutate(speech = paste0(speechdate, ": ", toupper(speaker), " ", speech)) %>%
  select(speech)

peel_speeches %>%
  slice_head(n = 5) %>%
  mutate(speech = str_trunc(speech, 220)) %>%
  knitr::kable(format = "latex", booktabs = TRUE,
               caption = "Peel - speeches containing “assemblies” (truncated)") %>%
  kableExtra::kable_styling(full_width = FALSE, latex_options = c("hold_position","scale_down")) %>%
  kableExtra::column_spec(1, width = "14cm")

kable(peel_top_by_count, format = "latex", booktabs = TRUE)

kable(peel_top_by_tfidf, format = "latex", booktabs = TRUE)

peel_sentences %>%
  slice_head(n = 12) %>%
  mutate(text = str_trunc(text, 160)) %>%
  select(sentence_id, year, text) %>%
  kable(format = "latex", booktabs = TRUE) %>% 
  column_spec(3, width = "10cm")
```

While Buckingham and Peel both support “freedom” as a goal, their
fundamental views on it are completely different. Peel’s top words
include broader parliamentary terms like “slave,” “slavery,” and
“house,” as well as more institutional and procedural words such as
“resolution,” “legislature,” “plan,” and “colony,” along with “freedom”
and “emancipation.” This reveals that Peel likely takes a more
governance-first approach, treating abolition as a constitutional and
administrative issue that must be carefully managed through legal
authority. That focus becomes even clearer when looking at his use of
“assemblies” in context, where he discusses the “West-Indian assemblies”
and the problem of how emancipation can be implemented through colonial
political structures rather than imposed abruptly.

In contrast, Buckingham uses terms like "wage," “pay,” and "labour" to
frame freedom as a relationship grounded in choice and wages. He uses
“apprenticeship” as a top word, showing how he criticized the
“apprenticeship/servitude” system as another form of slavery, while
unexpected words like "sugar" and "indies" show how he drew direct
comparisons between colonial labor and British industry - actively
challenging race-based assumptions of inferiority by demanding, "Let the
sugar-cultivators of the colonies be as free to choose their occupations
as the steel-grinders of Sheffield." Buckingham argues that if
“labourers were free to choose whichever they preferred,” the right
incentives would guide them to different kinds of work. He believes that
immediate emancipation would allow former slaves to earn good wages,
build wealth, and participate in business on their own terms.

The key difference is that Buckingham focuses on moral urgency and labor
rights, while Peel is more concerned with government, law, and a
controlled process where “freedom” is given through “resolutions,” legal
authority, and “plans” for colonial reform.

## 4. Ask AI to tell you about the Abolition of Slavery, using a prompt that requires it to use published sources only and not hallucinate. Then read a Wikipedia article on the Abolition of Slavery. Compare both to what you learned about the Abolition of Slavery in the introduction to this chapter. Which did you find most digestible? Which is most based in fact? Which do you trust the most?

Prompt I used: "Explain the abolition of slavery in the British Empire
focusing on 1833–1838. Use published sources only. For each paragraph,
include at least one citation in the format (Author, Title, Year) or a
reputable archive/government site. Do not cite Wikipedia. If you cannot
support a claim, write ‘I don’t know’ and do not guess. Cover: the
Slavery Abolition Act (1833), the apprenticeship system, compensation,
and what changed in 1838."

**AI Output:** "*Parliament’s **Slavery Abolition Act** received royal
assent on **28 August 1833** and declared that people “holden in
slavery” in British colonies would become free on **1 August 1834**,
making slavery “utterly and for ever abolished and declared unlawful” in
those colonies. (UK Government, Slavery Abolition Act 1833, 1833,
(<https://www.legislation.gov.uk/ukpga/Will4/3-4/73/contents>)*

*In practice, the Act paired emancipation with an **“apprenticeship”**
transition: many formerly enslaved people were legally reclassified as
apprentices and required to continue laboring under their former owners’
authority. In the 1833 parliamentary settlement, the apprenticeship term
was set at **four years for non-prædial** workers and **six years for
prædial** (plantation/agricultural) workers. (UK Parliament, Hansard:
Abolition Of Negro Apprenticeship [Commons debate], 1838,
(<https://hansard.parliament.uk/commons/1838-03-29/debates/b540ba24-695f-4827-95de-b4dfa1fcf2ba/AbolitionOfNegroApprenticesship>)*

*The Act also created a **compensation** settlement for slaveholders - a
fund of **£20 million** to compensate those previously “entitled to the
services” of enslaved people, while the formerly enslaved received **no
compensation**. The compensation process generated extensive claims
records and is treated by historians as a major state-financed payout to
slave-owners. (Draper, The Price of Emancipation, 2010,
<https://assets.cambridge.org/97805211/15254/frontmatter/9780521115254_frontmatter.pdf>)*

*By **1838**, the key change was that **apprenticeship ended early**:
under metropolitan pressure, colonial legislatures abolished
apprenticeship and **full freedom** was granted on **1 August 1838**,
ending the legal status that bound ex-slaves to compulsory labor as
“apprentices.” (UK Parliament, The West Indian colonies and
emancipation, n.d.,
(<https://www.parliament.uk/about/living-heritage/evolutionofparliament/legislativescrutiny/parliament-and-empire/parliament-and-the-american-colonies-before-1765/the-west-indian-colonies-and-emancipation>)"*

##### The Wikipedia article I read: (<https://en.wikipedia.org/wiki/Slavery_Abolition_Act_1833>)

Compared to ChatGPT's explanation, Wikipedia's article on the Slavery
Abolition Act was much more comprehensive. The AI summary was more
digestible because it correctly identified the core timeline (August
1833 to August 1838) and immediate details. However, the AI summary
never gave the context about what had happened before 1833. Wikipedia
had a much more thorough background, mentioning the 1772 Somerset case,
widely interpreted as limiting slavery under English law, and an active
anti-slavery movement in 1783. Most importantly, while the AI summary had
correctly mentioned the "apprenticeship" system, it didn't tell me the
true brutal reality behind it- only slaves under age 6 were actually
free, and everyone aged six and older was forced into continued unpaid
labor for a while after the Act passed. Ultimately, the AI response is
useful for a quick surface-level overview, given that it follows
citation rules, but the Wikipedia article and the chapter introduction
seem far more trustworthy for capturing the actual historical realities
and the lived experiences of the enslaved.

## 5. Try using AI to summarize what the speakers said, how they converge and how they differ. Cut and paste the top_slavery_speakers_1833 dataset into an AI prompt. What are the results? Did the AI highlight any of the most distinctive words, or did it move straight to summary? If an AI performs a summary accurately but doesn’t highlight the speakers’ most distinctive words, what would you be missing?

```

write_csv(top_slavery_speakers_1833, "top_slavery_speakers_data.csv")

```

Prompt: “I’m pasting the top_slavery_speakers_1833 dataset here. First,
write a 3–5 sentence summary of each speaker’s position. Second, list 5
exact words that are unusually characteristic of each speaker (quote the
exact words from the pasted text). Third, give 2 ways the speakers
converge and 2 ways they differ. Do not add outside facts.”

##### AI output:

*Mr. Fowell Buxton: He pushes for “substantial freedom” and treats
emancipation as urgent, warning that delay or half-measures risk unrest.
He rejects “apprenticeship” as basically slavery-by-another-name,
arguing there is “no medium between the system of the whip and a system
of wages.” He is willing to vote “compensation,” but only if Parliament
is truly “purchasing” freedom rather than paying while keeping coercion.
Characteristic words: “wages”, “apprenticeship”, “planters”, “negroes”,
“contract”.*

*The Duke of Wellington: He frames immediate emancipation as dangerous,
repeatedly emphasizing “security” and “serious evils” for colonies and
for the enslaved themselves. He stresses “rights of private property,”
supports “compensation,” and urges Parliament to “proceed slowly,” with
attention to colonial “legislatures” (especially “Jamaica”). His
position is cautious, privileging stability, legality, and a controlled
transition. Characteristic words: “legislatures”, “colonists”,
“compensation”, “petitioners”, “slowly”.*

*Mr. Buckingham: He argues for immediate emancipation “without any
further delay,” treating freedom as a claim of “justice” and “rights.”
He attacks “apprenticeship” as continued “forced subjection,” and frames
postponement as morally and politically indefensible. He accepts
compensation in principle but wants it framed as indemnifying a “loss,”
and he proposes policy levers like “tax” and trade preferences to shift
incentives away from “slave-labour.” Characteristic words: “bondage”,
“rights”, “tax”, “cultivation”, “delay”.*

*Sir Robert Peel: He accepts liberty as a principle but keeps returning
to the need to fit it to “the state of society in the colonies,” and he
is skeptical of immediate emancipation even if “demanded.” He emphasizes
practical governance: “regulations,” enforcement, evidence in courts,
and minimizing risk to “life and property.” He supports compensation and
a “progressive” path, warning against unleashing freedom without
preparation. Characteristic words: “stimulus”, “regulations”, “liberty”,
“canning”, “domingo”.*

*Mr. Stanley: He defends a government “plan” and rejects “immediate
emancipation” as “ruinous,” dangerous, and likely to provoke
“insurrection.” He argues for “apprenticeship” as a “probationary”
interval between “absolute slavery” and “absolute freedom,” and he
stresses economic continuity (especially “sugar” and “revenue”). He also
worries about uneven change across colonies, naming places like
“Trinidad” and “Jamaica.” Characteristic words: “probationary”,
“compulsory”, “ruinous”, “trinidad”, “sugar”.*

*Convergence: (1) All treat slavery’s end as on the agenda and debate
how to structure the transition (freedom, labor, law). (2) All
foreground consequences, whether for “security,” “insurrection,”
“revenue,” or the conditions of labor after slavery.*

*Differences: (1) Speed and form: Buckingham presses immediate freedom;
Wellington, Peel, and Stanley argue for gradualism and safeguards;
Buxton demands real freedom but rejects apprenticeship. (2) Moral-rights
vs. property-governance emphasis: Buckingham leans on “rights” and
justice; Wellington and Peel repeatedly center property, legality, and
stability; Stanley centers administrative plan + economic continuity;
Buxton centers wages-based labor and the integrity of what Parliament is
paying for.*

When I fed the raw dataset into an AI with strict prompt engineering, it
successfully generated a readable thematic summary of the main divisions
in the debate: immediate emancipation vs gradual transition, and
rights-based arguments vs those focused on property or governance.
 However, comparing the AI’s “characteristic words” to our earlier
tf-idf model reveals a critical discrepancy: ChatGPT did identify many
key terms, such as “wages” for Buxton and “legislatures” for Wellington,
but it also replaced many of the most statistically unique words with
more general labels. For example, it chose “stimulus” and “regulations”
as Peel’s main words, but missed his most distinctive tf-idf terms, such
as “resolution,” “colony,” and “plan.” If an analyst relies solely on an
AI summary without running the underlying calculations and using
validation techniques, as discussed in our readings on The Dangerous Art
of Text Mining, we might lose the evidence linking the interpretation to
measurable language patterns in the data. Distinctive words help show
which ideas each speaker returns to and how their arguments are
structured. Without those verified anchor terms, you miss the rhetorical
and conceptual texture of the debate and risk settling for an AI’s
thematic approximations rather than claims you can justify directly from
the text.

## 6. In this chapter, we saw how tf-idf can help us to distinguish the words of different speakers. Compare these results to the analysis in Chapter 3, when we simply asked who spoke the most. What is added by investigating the speakers’ words that you couldn’t have known by just modeling word count? What is added by modeling word count that you couldn’t have known by just investigating words?

Word count analysis reveals who dominated the floor and which voices had
the biggest chance to influence the debate. It gives us a sense of who
set the agenda, got attention, and held power in Parliament. However,
word count alone cannot tell us what speakers were actually doing with
language - two MPs can speak the same amount while advancing
fundamentally different arguments.

tf-idf fills in what word counts miss by showing how speakers frame
issues differently, even when discussing the same topic. It brings out
the unique ideas each speaker uses that word count alone cannot show.
For example, Buckingham emphasizes “wages” and labor freedom, while Peel
focuses on “assemblies,” “legislature,” and a planned approach to
colonial emancipation. In short, tf-idf helps us see the strategies and
disagreements in how speakers use language, what “freedom” means to
them, which policies they care about, and how they defend their views.

At the same time, word count gives us something tf-idf cannot: a sense
of who holds political power and who is likely to have an impact. Just
because someone uses distinctive language does not mean they are
influential if that speaker rarely takes the floor. On the other hand,
high-frequency speakers may shape outcomes even with less distinctive
vocabularies. Used together, word count shows us where power lies, while
tf-idf reveals the substance and structure of competing arguments.

## AI Use Statement

I used AI tools in the following ways:

**Tool:** Chatgpt

**Purpose:** I used ChatGPT for questions 3-6 and for vector allocation
issues ("vec_size") I encountered when initially loading data.

**Example Prompts:** The prompts used for text generation are included
in questions 4 and 5.

**Independent Work:** I wrote and ran the R code for data joining,
tokenizing, filtering, and tf-idf calculations in RStudio. I also did
the historical interpretation by checking the printed tables and context
sentences, identifying distinct conceptual frameworks (for example,
Buckingham’s “wage”-based labor freedom versus Peel’s focus on colonial
“assemblies” and governance), and writing my final interpretations.
